### 2018-03-26 02_scripts/01_hyperspectral_indices/01_hs-rename-and-subset.Rmd
add files

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

# Load packages

```{r}
library(sf)
library(raster)
library(magrittr)
library(here)
library(purrr)
library(dplyr)
```

# Read all atmoshperic corrected hyperspectral images

```{r}
map(list.files("/data/patrick/raw/hyperspectral/ATM", 
                    full.names = TRUE, pattern = ".tif$"),
    ~ brick(.x)) -> data_atm

all.plots <- st_read("/home/shares/data/LIFE/mod/plots/location/all.plots.shp") %>% mutate(Name = as.character(Name))
```

# Read plot locations

```{r}
st_read("/data/patrick/mod/plots/location/all.plots.shp") %>% 
  mutate(Name = as.character(Name)) -> plot_locations
```

# Load plots by name and perform renaming and subsetting for each

Names are extracted from the first three digits in the file name
e.g. B101 -> Plot 'Busturi- axpe' (see README for LUT)

1. Change plot shp CRS to HS CRS (faster than changing the CRS of large raster)
2. Crop HS by plot shp
3. Apply UTM Zone 30 CRS
4. Save cropped raster

## Busturi- axpe (Code B101)
```{r}
busturi1 <- raster::brick(files[[1]])

all.plots %>% 
  filter(Name == "a busturi-axpe") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

busturi1 <- raster::crop(busturi1, plot.shp)
busturi1 <- raster::mask(busturi1, plot.shp)
busturi1 <- raster::projectRaster(busturi1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(busturi1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/busturi.tif",
                    overwrite = TRUE)
```

## Laukiz 1, Laukiz 2, Laukiz 3 (Code B103)

```{r}
## Laukiz I
laukiz1 <- raster::brick(files[[2]])

all.plots %>% 
  filter(Name == "Laukiz I") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laukiz1 <- raster::crop(laukiz1, plot.shp)
laukiz1 <- raster::mask(laukiz1, plot.shp)
laukiz1 <- raster::projectRaster(laukiz1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laukiz1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz1.tif",
                    progress = "text", overwrite = TRUE)

## Laukiz II
laukiz2 <- raster::brick(files[[2]])

all.plots %>% 
  filter(Name == "Laukiz II") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laukiz2 <- raster::crop(laukiz2, plot.shp)
laukiz2 <- raster::mask(laukiz2, plot.shp)
laukiz2 <- raster::projectRaster(laukiz2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laukiz2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz2.tif",
                    progress = "text", overwrite = TRUE)

## Laukiz III
laukiz3 <- raster::brick(files[[2]])

all.plots %>% 
  filter(Name == "Laukiz III") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laukiz3 <- raster::crop(laukiz3, plot.shp)
laukiz3 <- raster::mask(laukiz3, plot.shp)
laukiz3 <- raster::projectRaster(laukiz3, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laukiz3, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laukiz3.tif",
                    progress = "text", overwrite = TRUE)

```

## Altube (Code B111 and B211)

```{r}
altube <- raster::brick(files[[3]])

all.plots %>% 
  filter(Name == "a altube") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

altube <- raster::crop(altube, plot.shp)
altube <- raster::mask(altube, plot.shp)
altube <- raster::projectRaster(altube, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(altube, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/altube.tif", 
                    progress = "text", overwrite = TRUE) 
```

## Barazar (Code B113)

```{r}
barazar <- raster::brick(files[[4]])

all.plots %>% 
  filter(Name == "a barazar") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

barazar <- raster::crop(barazar, plot.shp)
barazar <- raster::mask(barazar, plot.shp)
barazar <- raster::projectRaster(barazar, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(barazar, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/barazar.tif",
                    progress = "text", overwrite = TRUE)
```

## Laricio defol. Urkiola (Code B114)

```{r}
Laricio.defol.urkiola <- raster::brick(files[[5]])

all.plots %>% 
  filter(Name == "Laricio defol. Urkiola.") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

Laricio.defol.urkiola <- raster::crop(Laricio.defol.urkiola, plot.shp)
Laricio.defol.urkiola <- raster::mask(Laricio.defol.urkiola, plot.shp)
Laricio.defol.urkiola <- raster::projectRaster(Laricio.defol.urkiola, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(Laricio.defol.urkiola, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio-defol-urkiola.tif",
                    progress = "text", overwrite = TRUE)
```

## Aguinaga pinaster radiata (Code B123)

```{r}
aguinaga.pinaster.radiata <- raster::brick(files[[6]])

all.plots %>% 
  filter(Name == "aguinaga pinaster radiata") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

aguinaga.pinaster.radiata <- raster::crop(aguinaga.pinaster.radiata, plot.shp)
aguinaga.pinaster.radiata <- raster::mask(aguinaga.pinaster.radiata, plot.shp)
aguinaga.pinaster.radiata <- raster::projectRaster(aguinaga.pinaster.radiata, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(aguinaga.pinaster.radiata, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/aguinaga-pinaster-radiata.tif",
                    progress = "text", overwrite = TRUE)
```

## Oiartzun (Code B124)

```{r}
oiartzun <- raster::brick(files[[7]])

all.plots %>% 
  filter(Name == "Oiartzun") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

oiartzun <- raster::crop(oiartzun, plot.shp)
oiartzun <- raster::mask(oiartzun, plot.shp)
oiartzun <- raster::projectRaster(oiartzun, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(oiartzun, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/oiartzun.tif",
                    progress = "text", overwrite = TRUE)
```

## Pinast. Arza (Code B204)

```{r}
pinast.arza <- raster::brick(files[[8]])

all.plots %>% 
  filter(Name == "Pinast. Arza") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

pinast.arza <- raster::crop(pinast.arza, plot.shp)
pinast.arza <- raster::mask(pinast.arza, plot.shp)
pinast.arza <- raster::projectRaster(pinast.arza, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(pinast.arza, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/pinast-arza.tif",
                    progress = "text", overwrite = TRUE) 
```

## Pinas. Santa Koloma (Code B205)

```{r}
pinas.santa.koloma <- raster::brick(files[[9]])

all.plots %>% 
  filter(Name == "Pinas. Santa Koloma") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

pinas.santa.koloma <- raster::crop(pinas.santa.koloma, plot.shp)
pinas.santa.koloma <- raster::mask(pinas.santa.koloma, plot.shp)
pinas.santa.koloma <- raster::projectRaster(pinas.santa.koloma, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(pinas.santa.koloma, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/pinas-santa-koloma.tif",
                    progress = "text", overwrite = TRUE)
```

## Rad. Menagaray 1 and Rad. Mengaray 2 (Code B206)

```{r}
## Rad. Menagaray 1
rad.menagaray.1 <- raster::brick(files[[10]])

all.plots %>% 
  filter(Name == "Rad. Menagaray 1") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

rad.menagaray.1 <- raster::crop(rad.menagaray.1, plot.shp)
rad.menagaray.1 <- raster::mask(rad.menagaray.1, plot.shp)
rad.menagaray.1 <- raster::projectRaster(rad.menagaray.1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(rad.menagaray.1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-menagaray-1.tif",
                    progress = "text", overwrite = TRUE)

## Rad. Menagaray 2
rad.menagaray.2 <- raster::brick(files[[10]])

all.plots %>% 
  filter(Name == "Rad. Menagaray 2") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

rad.menagaray.2 <- raster::crop(rad.menagaray.2, plot.shp)
rad.menagaray.2 <- raster::mask(rad.menagaray.2, plot.shp)
rad.menagaray.2 <- raster::projectRaster(rad.menagaray.2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(rad.menagaray.2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-menagaray-2.tif",
                    progress = "text", overwrite = TRUE) 
```

## Ayala_aiara / Luiando (Code B207)

```{r}
ayala.aiara <- raster::brick(files[[11]])

all.plots %>% 
  filter(Name == "Ayala/Aiara") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

ayala.aiara <- raster::crop(ayala.aiara, plot.shp)
ayala.aiara <- raster::mask(ayala.aiara, plot.shp)
ayala.aiara <- raster::projectRaster(ayala.aiara, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(ayala.aiara, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/ayala-aiara.tif",
                    progress = "text", overwrite = TRUE)
```

## Silvest. Nograro (Code B208)

```{r}
silvest.nograro <- raster::brick(files[[12]])

all.plots %>% 
  filter(Name == "Silvest. Nograro") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

silvest.nograro <- raster::crop(silvest.nograro, plot.shp)
silvest.nograro <- raster::mask(silvest.nograro, plot.shp)
silvest.nograro <- raster::projectRaster(silvest.nograro, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(silvest.nograro, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/silvest-nograro.tif",
                    progress = "text", overwrite = TRUE)
```

## Caranca (Code B209)

```{r}
caranca <- raster::brick(files[[13]])

all.plots %>% 
  filter(Name == "a caranca") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

caranca <- raster::crop(caranca, plot.shp)
caranca <- raster::mask(caranca, plot.shp)
caranca <- raster::projectRaster(caranca, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(caranca, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/caranca.tif",
                    progress = "text", overwrite = TRUE)
```

## Picea Abornicano (Code B210)

```{r}
picea.abornicano <- raster::brick(files[[14]])

all.plots %>% 
  filter(Name == "Picea Abornicano") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

picea.abornicano <- raster::crop(picea.abornicano, plot.shp)
picea.abornicano <- raster::mask(picea.abornicano, plot.shp)
picea.abornicano <- raster::projectRaster(picea.abornicano, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(picea.abornicano, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/picea-abornicano.tif",
                    progress = "text", overwrite = TRUE)
```

## Douglas 1 and Douglas 2 and Rad. Altube (Code B212)

Code 211 bzw. file[[15]] == Altube

```{r}
## douglas 1
douglas.1 <- raster::brick(files[[16]])

all.plots %>% 
  filter(Name == "douglas 1") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.1 <- raster::crop(douglas.1, plot.shp)
douglas.1 <- raster::mask(douglas.1, plot.shp)
douglas.1 <- raster::projectRaster(douglas.1, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.1, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas.1.tif",
                    progress = "text", overwrite = TRUE)

## douglas 2
douglas.2 <- raster::brick(files[[16]])

all.plots %>% 
  filter(Name == "douglas 2") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.2 <- raster::crop(douglas.2, plot.shp)
douglas.2 <- raster::mask(douglas.2, plot.shp)
douglas.2 <- raster::projectRaster(douglas.2, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.2, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-2.tif",
                    progress = "text", overwrite = TRUE)

## rad. altube
rad.altube <- raster::brick(files[[16]])

all.plots %>% 
  filter(Name == "Rad. Altube") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

rad.altube <- raster::crop(rad.altube, plot.shp)
rad.altube <- raster::mask(rad.altube, plot.shp)
rad.altube <- raster::projectRaster(rad.altube, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(rad.altube, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/rad-altube.tif",
                    progress = "text", overwrite = TRUE)
```

## Otxandio and Laricio defol. Olaeta (Code B215)

```{r}
## otxandio
otxandio <- raster::brick(files[[17]])

all.plots %>% 
  filter(Name == "a otxandio") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

otxandio <- raster::crop(otxandio, plot.shp)
otxandio <- raster::mask(otxandio, plot.shp)
otxandio <- raster::projectRaster(otxandio, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(otxandio, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/otxandio.tif",
                    progress = "text", overwrite = TRUE)

## Laricio defol. Olaeta
laricio.defol.olaeta <- raster::brick(files[[17]])

all.plots %>% 
  filter(Name == "Laricio defol. Olaeta. pol") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laricio.defol.olaeta <- raster::crop(laricio.defol.olaeta, plot.shp)
laricio.defol.olaeta <- raster::mask(laricio.defol.olaeta, plot.shp)
laricio.defol.olaeta <- raster::projectRaster(laricio.defol.olaeta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laricio.defol.olaeta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio-defol-olaeta.tif",
                    progress = "text", overwrite = TRUE)
```

## Douglas Legutio (Code B317 and B318)
Here B317 is taken
```{r}
douglas.legutio <- raster::brick(files[[18]])

all.plots %>% 
  filter(Name == "Douglas Legutio") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.legutio <- raster::crop(douglas.legutio, plot.shp)
douglas.legutio <- raster::mask(douglas.legutio, plot.shp)
douglas.legutio <- raster::projectRaster(douglas.legutio, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.legutio, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-legutio.tif",
                    progress = "text", overwrite = TRUE)
```
## Laricio Gatzaga (Code B319)

```{r}
laricio.gatzaga <- raster::brick(files[[20]])

all.plots %>% 
  filter(Name == "Laricio defol.Gatzaga") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

laricio.gatzaga <- raster::crop(laricio.gatzaga, plot.shp)
laricio.gatzaga <- raster::mask(laricio.gatzaga, plot.shp)
laricio.gatzaga <- raster::projectRaster(laricio.gatzaga, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(laricio.gatzaga, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/laricio.gatzaga.tif",
                    progress = "text", overwrite = TRUE)
```
## Azaceta (Code B320)

```{r}
azaceta <- raster::brick(files[[21]])

all.plots %>% 
  filter(Name == "a azaceta") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

azaceta <- raster::crop(azaceta, plot.shp)
azaceta <- raster::mask(azaceta, plot.shp)
azaceta <- raster::projectRaster(azaceta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(azaceta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/azaceta.tif",
                    progress = "text", overwrite = TRUE)
```
## Gipuzkoa and Sequoia Legorreta (Code B321)

```{r}
## gipuzkoa
gipuzkoa <- raster::brick(files[[22]])

all.plots %>% 
  filter(Name == "a gipuzkoa") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

gipuzkoa <- raster::crop(gipuzkoa, plot.shp)
gipuzkoa <- raster::mask(gipuzkoa, plot.shp)
gipuzkoa <- raster::projectRaster(gipuzkoa, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(gipuzkoa, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/gipuzkoa.tif",
                    progress = "text", overwrite = TRUE)

## Sequoia Legorreta
sequoia.legorreta <- raster::brick(files[[22]])

all.plots %>% 
  filter(Name == "Sequoia Legorreta") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

sequoia.legorreta <- raster::crop(sequoia.legorreta, plot.shp)
sequoia.legorreta <- raster::mask(sequoia.legorreta, plot.shp)
sequoia.legorreta <- raster::projectRaster(sequoia.legorreta, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(sequoia.legorreta, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/sequoia.legorreta.tif",
                    progress = "text", overwrite = TRUE)
```
## Douglas Albiztur (Code B322)

```{r}
douglas.albiztur <- raster::brick(files[[23]])

all.plots %>% 
  filter(Name == "Douglas Albiztur") %>% 
  st_transform("+proj=utm +zone=30 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs") %>% 
  as("Spatial") -> plot.shp

douglas.albiztur <- raster::crop(douglas.albiztur, plot.shp)
douglas.albiztur <- raster::mask(douglas.albiztur, plot.shp)
douglas.albiztur <- raster::projectRaster(douglas.albiztur, crs = "+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs")

raster::writeRaster(douglas.albiztur, "/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/douglas-albiztur.tif",
                    progress = "text", overwrite = TRUE)
```