### 2018-03-26 02_scripts/01_hyperspectral_indices/02_hyperspectral-ndvi-raster.Rmd

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

```{r}
pacman::p_load(raster, hsdar, here, pbapply, parallel)
```

# specify wavelength

```{r}
wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")
```

# read rasters (cropped by plot)

## list files

```{r}
files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
                    pattern = ".tif$", full.names = TRUE)
```

## read in files as "RasterBrick"

```{r}
spectral_data <- lapply(files, function(x) brick(x))
```

# make class "HyperSpecRaster"
```{r}
hyperspecs <- lapply(spectral_data, function(x) hsdar::HyperSpecRaster(x, wavelength))
```

# account for use on local laptop or server
```{r}
if (parallel::detectCores() > 10) {
  cores <- 28
} else (cores <- 1)
```

# calculate NDVI

assign to object to prevent list printing to console ### use 'NDVI2' as formula 
```{r}
pboptions(style = 1, type = "timer")
tmp <- pblapply(hyperspecs, function(x) 
  vegindex(x, "NDVI2", filename = paste0("/home/shares/data/LIFE/mod/hyperspectral/ndvi/", "ndvi2-",
                                         basename(x@file@name)),
           bnames = "NDVI"), cl = cores)
```