### 2018-03-26 02_scripts/01_hyperspectral_indices/03_hyperspectral-ndvi-analysis.Rmd

---
title: "NDVI analysis of hyperspectral data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
# Data & packages

Loads list of 28 cropped (to respective plot) atmospherical corrected hyperspectral images.  
See R script `hs.ndvi.R` for multicore creation.

Move to chunk 'hsdar package' if file exists!
```{r}
pacman::p_load(rgdal, raster, mapview, hsdar, rio, ggplot2, colorRamps)

if (file.exists("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/all-cropped.rds")) {
  files <- as.list(rio::import("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/all-cropped.rds"))
} else {
  files <- list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped/", 
                    pattern = ".tif$", full.names = TRUE)

  files1 <- lapply(files, function(x) raster::brick(x))
}
```

Define 'Get values row by row' (vignette `raster::functions`)
* x = multy layer Raster file (applying `brick()`)
* a = value to add to existing value
```{r}
source(here("R/helper.funs/get.values.row.by.row.R"))
```

# Get values from all raster files (bricks!) using `lapply`
```{r}
# files_atm <- lapply(files_atm, function(x) f8(x, 0, progress = "text"))
# saveRDS(files_atm, here("R/tmp/files_atm_with_values.rda")
files_atm <- readRDS("/home/patrick/PhD/R/tmp/files_atm_with_values.rda")
```

Get wavelengths out of README table (http://datascienceplus.com/extracting-tables-from-pdfs-in-r-using-the-tabulizer-package/)

Does not work because wavelength table on page 10 of README is not a table but an image...
```{r}
# devtools::install_github("ropenscilabs/tabulizer")
# devtools::install_github("ropenscilabs/tabulizerjars")

# library(tabulizer)

# tabs <- extract_tables("/home/shares/data/original/hyperspectral/0_README_spanish.pdf", pages = 10)
```

# hsdar package
```{r}
# specify wavelength of each band
wavelength <- c(404.08, 408.5, 412.92, 417.36, 421.81, 426.27, 430.73, 435.20, 
                439.69, 444.18, 448.68, 453.18, 457.69, 462.22, 466.75, 471.29,
                475.83, 480.39, 484.95, 489.52, 494.09, 498.68, 503.26, 507.86,
                512.47, 517.08, 521.70, 526.32, 530.95, 535.58, 540.23, 544.88,
                549.54, 554.20, 558.86, 563.54, 568.22, 572.90, 577.60, 582.29,
                586.99, 591.70, 596.41, 601.13, 605.85, 610.58, 615.31, 620.05,
                624.79, 629.54, 634.29, 639.04, 643.80, 648.56, 653.33, 658.10,
                662.88, 667.66, 672.44, 677.23, 682.02, 686.81, 691.60, 696.40,
                701.21, 706.01, 710.82, 715.64, 720.45, 725.27, 730.09, 734.91,
                739.73, 744.56, 749.39, 754.22, 759.05, 763.89, 768.72, 773.56, 
                778.40, 783.24, 788.08, 792.93, 797.77, 802.62, 807.47, 812.32,
                817.17, 822.02, 826.87, 831.72, 836.57, 841.42, 846.28, 851.13,
                855.98, 860.83, 865.69, 870.54, 875.39, 880.24, 885.09, 889.94,
                894.79, 899.64, 904.49, 909.34, 914.18, 919.03, 923.87, 928.71,
                933.55, 938.39, 943.23, 948.07, 952.90, 957.73, 962.56, 967.39,
                972.22, 977.04, 981.87, 986.68, 991.50, 996.31)
```

Read in raster data either using `rgdal` or `raster` package. 
`raster` seems to be prefered but `rgdal` is a lot faster in reading values

```{r}
## create HyperSpecRaster file
# ra <- hsdar::HyperSpecRaster(files_atm[[1]], wavelength)
# tr <- blockSize(ra)

##  create SpecLib file
# spectral_data1 <- f8(files1[[1]], 0, progress = "text")

# speclib <- hsdar::speclib(spectral_data, wavelength)
```

plot spectral signature
```{r}
# plot(spectral_data, main = "Default Plot")

# plot(spectral_data, FUN = "median") # not working
```

Visualize NDVI
```{r}
mapview::mapview(spectral_data1, zcol = "ndvi", maxpixels = 4664960, na.color = "white")
```

# NDVI analysis

## Read in files

1. List files
2. Read in using `rgdal::readGDAL`
3. Get filenames
4. Rename `band1` to `NDVI` and label SpGrDaFr with plot name
4. Get pixel count of each file
5. Create DF with one NDVI column and a factor column with filenames
```{r}
files.list <- list.files("/home/shares/data/LIFE/mod/hyperspectral/ndvi", 
                    pattern = "ndvi2.*\\.tif", full.names = TRUE)

files.ndvi <- lapply(files.list, function(x) rgdal::readGDAL(x, silent = TRUE))
# rename band name to 'NDVI'
for (i in seq_along(files.ndvi)) {
  colnames(files.ndvi[[i]]@data) <- "NDVI"

}

names.file <- sapply(files.list, function(x) basename(x))
# label SpGrDaFr with plot name
names(files.ndvi) <- names.file

pixel.count <- sapply(files.ndvi, function(x) length(x@data$NDVI))

ndvi.all <- c()
ndvi.all <- unlist(lapply(files.ndvi, function(x) append(ndvi.all, x@data$NDVI)))

plot.names <- c()
for (i in seq_along(names.file)) {
  plot.names <- append(plot.names, rep(basename(names.file[i]), pixel.count[i]))
}

df <- data.frame(NDVI = ndvi.all,
                 plot = as.factor(plot.names))

# remove NA rows
df <- df[complete.cases(df),]
```

## Plot NDVI histograms

```{r}
ndvi.histograms <- ggplot(df, aes(NDVI)) +
  #geom_histogram() + 
  # normalize y-axis of each facet http://stackoverflow.com/questions/16339204/normalizing-faceted-histograms-separately-in-ggplot2
  geom_histogram(aes(y = (..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
  facet_wrap(~ plot) + 
  ylab("Percentage pixel count") + 
  xlab("NDVI (0.25 to 1)") + 
  scale_x_continuous(limits = c(0.25, 1)) + 
  cowplot::background_grid()
```

## Count pixels per plot (size 1m^2)
```{r}
table(df$plot)
```

## Quantiles

Biased towards low values (non-forest pixels)
```{r}
stats::quantile(df$NDVI)
```

```{r}
boxplot(df$NDVI)
```

```{r}
color <- colorRamps::green2red(2)
rasterVis::levelplot(files.ndvi[[4]])
mapview::mapView(as.raster(files.ndvi[[4]]), zcol = "NDVI", na.color = "blue", color = color)
```

## plot NDVI rasters
```{r}
# convert to raster layer for plotting
files.ndvi.raster <- lapply(files.ndvi, function(x) raster::raster(x))
#set names
for (i in seq_along(files.ndvi.raster)) {
  files.ndvi.raster[[i]]@file@name <- names(files.ndvi)[i]
}

# using mapview
list.plots <- lapply(files.ndvi.raster, function(x) mapview::mapview(x, use.layer.names = TRUE, na.color = "white"))
# using rasterVis::levelplot
list.plots <- lapply(files.ndvi.raster, function(x) rasterVis::levelplot(x,
                                                                         main = x@file@name,
                                                                         margin = FALSE,
                                                                         pretty = TRUE,
                                                                         col.regions = rev(colorRamps::green2red(400)), 
                                                                         at = seq(0.5, 1, len = 101))
)

gridExtra::grid.arrange(list.plots[[1]], list.plots[[2]], list.plots[[3]], 
                                        list.plots[[4]], list.plots[[5]], list.plots[[6]],
                                        list.plots[[7]], list.plots[[8]], list.plots[[9]],
                                        ncol = 3)

gridExtra::grid.arrange(list.plots[[10]], list.plots[[11]], list.plots[[12]],
                                        list.plots[[13]], list.plots[[14]], list.plots[[15]],
                                        list.plots[[16]], list.plots[[17]], list.plots[[18]],
                                        ncol = 3)

gridExtra::grid.arrange(list.plots[[19]], list.plots[[20]], list.plots[[21]],
                                        list.plots[[22]], list.plots[[23]], list.plots[[24]], 
                                        list.plots[[25]], list.plots[[26]], list.plots[[27]], 
                                        list.plots[[28]],
                                        ncol = 3)

```

## save NDVI raster plots

```{r}
png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.1.png", width = 16.51, height = 12.07, units = "in", res = 300)
gridExtra::grid.arrange(list.plots[[1]], list.plots[[2]], list.plots[[3]], 
                                        list.plots[[4]], list.plots[[5]], list.plots[[6]],
                                        list.plots[[7]], list.plots[[8]], list.plots[[9]],
                                        ncol = 3)
dev.off()

png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.2.png", width = 16.51, height = 12.07, units = "in", res = 300)
gridExtra::grid.arrange(list.plots[[10]], list.plots[[11]], list.plots[[12]],
                                        list.plots[[13]], list.plots[[14]], list.plots[[15]],
                                        list.plots[[16]], list.plots[[17]], list.plots[[18]],
                                        ncol = 3)
dev.off()

png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi2.plots.3.png", width = 16.51, height = 12.07, units = "in", res = 300)
gridExtra::grid.arrange(list.plots[[19]], list.plots[[20]], list.plots[[21]],
                                        list.plots[[22]], list.plots[[23]], list.plots[[24]], 
                                        list.plots[[25]], list.plots[[26]], list.plots[[27]], 
                                        list.plots[[28]],
                                        ncol = 3)
dev.off()
```

## save ndvi histograms
```{r}
png("/home/patrick/PhD/figs/hyperspectral/ndvi/ndvi.histograms.normalized.png", width = 16.51, height = 12.07, units = "in", res = 300)
ndvi.histograms
dev.off()
```