### 2018-03-26 02_scripts/01_hyperspectral_indices/05_narrow-band-indices.Rmd

---
title: "Calculate all possible narrow band indices"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

**Attention**

This script uses the function `nbi_raster()` which is a custom function and only available from the developtment version of `hsdar` on Github: `devtools::install_github("pat-s/hsdar")`.

```{r}
pacman::p_load(raster, hsdar, pbapply, parallel, tidyverse, here, magrittr)
```

# Load wavelength and rasters

```{r}
wavelength <- readRDS("/home/shares/data/LIFE/mod/hyperspectral/wavelength.Rds")

list.files("/home/shares/data/LIFE/mod/hyperspectral/atm-cropped", 
           pattern = ".tif$", full.names = TRUE) %>%
  lapply(function(x) brick(x)) %>% 
  lapply(function(x) HyperSpecRaster(x, wavelength)) -> hyperspecs
```

# account for use on local laptop or server
```{r}
if (detectCores() > 10) {
  cores <- 28
} else (cores <- 1)
```

# do calculation of narrow band indices

Band names will be calculated automatically by function `nbi_raster`

Takes 190 GB RAM on server when being parallalized.

Runtime per plot between 5 - 27 minutes.

Largest file: 24 GB

```{r}
## Calculate NBI
# hyperspecs <- list(hyperspecs[[1]])

hyperspecs %>% 
  pblapply(function(x) 
    nbi_raster(x, filename = gsub(".tif", ".grd", 
                                  "/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/",
                                         "narrow-band-indices-", basename(x@file@name))), cl = cores) 
```