### 2018-03-26 02_scripts/01_hyperspectral_indices/06_extract-indices-to-points.Rmd

---
title: "Extract indices to points"
output: html_notebook
editor_options: 
chunk_output_type: inline
---

# Information about this script

This script accounts for either local (laptop) or server (MCCOY) runs. Path is adjusted using object `prefix`

Since this is an `.Rmd` file it cannot simply be called using `source()` or `Rscript` (server).

Instead, please use the [custom function `runAllChunks()`](http://stackoverflow.com/a/24754637/4185785) inside an R console with the `.Rmd` file as input. 
It is recommended to put this function into your .Rprofile to have it available whenever you start an R session.

```{r}
pacman::p_load(raster, sf, here, magrittr, tidyverse, pbapply, parallel)
```

# Laukiz1

## Load points

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/laukiz1.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> laukiz1_points
```

## Load rasters

```{r}
laukiz1_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz1")
```

```{r}
laukiz1_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz1")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(laukiz1_veg_indices, laukiz1_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Nested parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(laukiz1_nbi, laukiz1_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = 5)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

laukiz1_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> laukiz1_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(laukiz1_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz1/all-indices-laukiz1.rda")
```

# Laukiz2

## Load points

```{r}
print("Starting Laukiz2")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/laukiz2.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> laukiz2_points
```

## Load rasters

```{r}
laukiz2_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-laukiz2")
```

```{r}
laukiz2_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-laukiz2")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(laukiz2_veg_indices, laukiz2_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(laukiz2_nbi, laukiz2_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = 5)
  #endCluster()
  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}

### local debugging code
# out_nbi <- raster::extract(laukiz2_nbi, laukiz2_points, buffer = 2, 
#                            fun = mean, df = TRUE)
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

laukiz2_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> laukiz2_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(laukiz2_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/laukiz2/all-indices-laukiz2.rda")
```

# Oiartzun

## Load points

```{r}
print("Starting oiartzun")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/oiartzun.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> oiartzun_points
```

## Load rasters

```{r}
oiartzun_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-oiartzun")
```

```{r}
oiartzun_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-oiartzun")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(oiartzun_veg_indices, oiartzun_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(oiartzun_nbi, oiartzun_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = cores)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

oiartzun_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> oiartzun_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(oiartzun_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/oiartzun/all-indices-oiartzun.rda")
```

# Luiando

## Load points

```{r}
print("Starting luiando")
```

```{r}
st_read("/home/shares/data/LIFE/mod/tree-per-tree/luiando/luiando.shp", 
        quiet = TRUE) %>% 
  as("Spatial") -> luiando_points
```

## Load rasters

```{r}
luiando_veg_indices <- brick("/home/shares/data/LIFE/mod/hyperspectral/all-veg-indices/all-veg-indices-ayala-aiara") 
```

```{r}
luiando_nbi <- brick("/home/shares/data/LIFE/mod/hyperspectral/narrow-band-indices/narrow-band-indices-ayala-aiara")
```

## Extract values to points

### Veg indices

Buffer size:

* 1m
* 2m
* 3m
* 4m
* 5m

```{r}
out_veg <- lapply(seq_along(1:5), function(x) 
  raster::extract(luiando_veg_indices, luiando_points, buffer = x, 
                  fun = mean, df = TRUE))
```

Label output names according to buffer size

```{r}
out_veg <- lapply(seq_along(out_veg), function(x) 
  setNames(out_veg[[x]], paste0("bf", x, "_", colnames(out_veg[[x]]))))
```

### NBI

This might take a while - nested parallelization here! 
Only outer parallelization: 1 hour

This can only be run on a server infrastructure! This code uses 8*5 = 40 + 5 = 45 cores!

NOTE 05 April: Neste parallelization causes some weird errors for Laukiz2 and does not really speed up the processing. 
I think we are better of using only outer parallelization here.

```{r}
if (detectCores() > 45 ) {

  Sys.time <- Sys.time()

  out_nbi <- pblapply(seq_along(1:5), function(x) {
    #beginCluster(8) # assigning 8 cores to each parallelized `extract` call -> 40 cores total because 8*5
    raster::extract(luiando_nbi, luiando_points, buffer = x, 
                    fun = mean, df = TRUE) } , cl = cores)

  print(Sys.time() - Sys.time)
} else {
  stop("please run this script on a server with more than 45 cores!")
}
```

### Bind to points

```{r}
# merge all data frames
df_out <- bind_cols(out_veg)

df_out1 <- bind_cols(out_nbi)

luiando_points %>% 
  st_as_sf() %>% 
  bind_cols(df_out) %>% 
  bind_cols(df_out1) -> luiando_points
```

## Save files

```{r}
# .rda file (data frame)
as.data.frame(luiando_points) %>% 
  saveRDS("/home/shares/data/LIFE/mod/tree-per-tree/luiando/all-indices-luiando.rda")
```