### 2018-03-26 02_scripts/02_preprocessing.Rmd

---
title: "Preprocessing of raw data"
author: "Patrick Schratz"
output:
  html_notebook:
    code_folding: none
    fig_width: 10
    highlight: haddock
    number_sections: yes
    toc: yes
---

# Load packages

```{r}
library(purrr)
library(here)
library(readr)
library(tibble)
library(glue)
library(raster)
```

# Read data

Reading in data from .txt files and exporting them as a Geopackage (.gpkg).

## Survey data

```{r}
names = c("laukiz1", "laukiz2", "luiando", "oiartzun")

map(names, 
    ~ as_tibble(read.table(file = glue("/data/patrick/mod/survey_data/2016/{.x}.txt"), 
                           sep = "\t", dec = ",", header = TRUE))) %>% 
  set_names(names) -> data_all
```

## Vegetation indices

```{r}
map(names,
    ~ brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/all-veg-indices-{.x}.gri"))) %>% 
  set_names(names) -> indices_all
```

## Narrow band indices

# Bind indices to response