### 2018-04-04 02_scripts/02_preprocessing.Rmd

---
title: "Preprocessing of raw data"
author: "Patrick Schratz"
output:
  html_notebook:
    code_folding: none
    fig_width: 10
    highlight: haddock
    number_sections: yes
    toc: yes
---

# Load packages

```{r}
library(purrr)
library(here)
library(readr)
library(tibble)
library(glue)
library(raster)
library(sf)
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
```

# Read data

Reading in data from .txt files and exporting them as a Geopackage (.gpkg).

## Survey data

```{r}
names = c("laukiz1", "laukiz2", "luiando", "oiartzun")

map(names, 
    ~ as_tibble(read.table(file = glue("/data/patrick/mod/survey_data/2016/{.x}.txt"), 
                           sep = "\t", dec = ",", header = TRUE,
                           stringsAsFactors = FALSE, 
                           na.strings=c("")))) %>% 
  set_names(names) %>% 
  map(~ rename_all(.x, tolower)) -> data_tbl
```

Laukiz2 has no coordinate info in the txt file. We need to read the shapefile and join them.

### Join Laukiz2 coords

```{r}
laukiz2_coords = st_read("/data/patrick/raw/survey_data/2016/laukiz2/arboles/Laukiz2_arboles_ETRS89.shp")

data_tbl$laukiz2 = left_join(data_tbl$laukiz2, laukiz2_coords, 
                             by = c("num.arbol" = "id_arbol"))
```

Drop observations with no coordinate information.

```{r}
data_tbl %<>% 
  map(~ drop_na(.x, c("x", "y")))
```

Coerce to `sf` objects.

```{r}
map(data_tbl, ~ st_as_sf(.x, coords = c("x", "y"))) -> data_sf
```

Select variables related to defoliation.

```{r}
data_sf %<>% 
  map(~ select(.x, contains("def")))
```

Laukiz2 needs some clean up:

"no" -> 0
"*" -> NA
Coerce from chr to int.
Calculate "deftot" which is the mean of all three parts (bottom, mid, top).

`rowSums` will thrown an error due to the "geometry" column so we need to add `drop = TRUE`.

```{r}
data_sf$laukiz2 %<>% 
  mutate(x..de.defoliacion.parte.media = 
           replace(.$x..de.defoliacion.parte.media, 
                   .$x..de.defoliacion.parte.media == "no", "0")) %>% 
  mutate(x..defoliacion.en.punta = 
           replace(.$x..defoliacion.en.punta, 
                   .$x..defoliacion.en.punta == "no", "0")) %>% 
  mutate(x..defoliacion.base = 
           replace(.$x..defoliacion.base,
                   .$x..defoliacion.base == "no", "0")) %>% 
  mutate(x..defoliacion.base = 
           replace(.$x..defoliacion.base,
                   .$x..defoliacion.base == "*", NA)) %>% 
  mutate_if(is.character, as.numeric) %>% 
  mutate(deftot = as.integer(rowMeans(.[, 1:3, drop = TRUE], na.rm = TRUE)))
```

## Vegetation indices

```{r}
map(names,
    ~ brick(glue("/data/patrick/mod/hyperspectral/all-veg-indices/all-veg-indices-{.x}.gri"))) %>% 
  set_names(names) -> veg_indices_all
```

## Narrow band indices

```{r}
map(names,
    ~ brick(glue("/data/patrick/mod/hyperspectral/narrow-band-indices/narrow-band-indices-{.x}.gri"))) %>% 
  set_names(names) -> nbi_indices_all
```

# Bind indices to response