### 2018-04-26 02_scripts/021_preprocessing.Rmd

---
title: "02-preprocessing"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r}
library(purrr)
library(here)
library(readr)
library(tibble)
library(glue)
library(raster)
library(sf)
library(dplyr)
library(tidyr)
library(magrittr)
library(stringr)
library(fs)
library(pbmcapply)
library(data.table)
```

# Read data

Reading in data from .txt files and exporting them as a Geopackage (.gpkg).

## Survey data

```{r}
names <- c("laukiz1", "laukiz2", "luiando", "oiartzun")

map(
  names,
  ~ as_tibble(read.table(
    file = glue("/data/patrick/mod/survey_data/2016/{.x}.txt"),
    sep = "\t", dec = ",", header = TRUE,
    stringsAsFactors = FALSE,
    na.strings = c("")
  ))
) %>%
  set_names(names) %>%
  map(~ rename_all(.x, tolower)) -> data_tbl
```

Laukiz2 has no coordinate info in the txt file. We need to read the shapefile and join them.

### Join Laukiz2 coords

```{r}
laukiz2_coords <- st_read("/data/patrick/raw/survey_data/2016/laukiz2/arboles/Laukiz2_arboles_ETRS89.shp")

data_tbl$laukiz2 <- left_join(data_tbl$laukiz2, laukiz2_coords,
  by = c("num.arbol" = "id_arbol")
)
```

Drop observations with no coordinate information.

```{r}
data_tbl %<>%
  map(~ drop_na(.x, c("x", "y")))
```

Coerce to `sf` objects.

```{r}
map(data_tbl, ~ st_as_sf(.x, coords = c("x", "y"))) -> data_sf
```

Select variables related to defoliation.

```{r}
data_sf %<>%
  map(~ dplyr::select(.x, contains("def")))
```

Laukiz2 needs some clean up:

"no" -> 0
"*" -> NA
Coerce from chr to int.
Calculate "deftot" which is the mean of all three parts (bottom, mid, top).

`rowSums` will thrown an error due to the "geometry" column so we need to add `drop = TRUE`.

```{r}
data_sf$laukiz2 %<>%
  mutate(
    x..de.defoliacion.parte.media =
      replace(
        .$x..de.defoliacion.parte.media,
        .$x..de.defoliacion.parte.media == "no", "0"
      )
  ) %>%
  mutate(
    x..defoliacion.en.punta =
      replace(
        .$x..defoliacion.en.punta,
        .$x..defoliacion.en.punta == "no", "0"
      )
  ) %>%
  mutate(
    x..defoliacion.base =
      replace(
        .$x..defoliacion.base,
        .$x..defoliacion.base == "no", "0"
      )
  ) %>%
  mutate(
    x..defoliacion.base =
      replace(
        .$x..defoliacion.base,
        .$x..defoliacion.base == "*", NA
      )
  ) %>%
  mutate_if(is.character, as.numeric) %>%
  mutate(deftot = as.integer(rowMeans(.[, 1:3, drop = TRUE], na.rm = TRUE))) %>%
  drop_na(deftot) %>%
  select(deftot, geometry)
```

# Write clean data as .gpkg

Writing to cifs mounts is currently not working so we have to copy and delete the files.

```{r}
# write
walk2(data_sf, seq_along(data_sf), ~ write_sf(.x, glue(here("{name}.gpkg"),
  name = names(data_sf[.y])
)))

# copy
walk(names(data_sf), ~ file_copy(glue(here("{.x}.gpkg")),
  glue("/data/patrick/mod/tree-per-tree/2016/{.x}/{.x}.gpkg"),
  overwrite = TRUE
))

# delete
walk(names(data_sf), ~ file_delete(glue(here("{.x}.gpkg"))))
```