### 2018-05-14 02_scripts/04_plot_characteristics.Rmd

---
title: "04-plot characteristics"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
devtools::install_github("tidyverse/ggplot2")
devtools::install_github("mlr-org/mlr@blocking")

library(dplyr)
library(tibble)
library(purrr)
library(magrittr)
library(data.table)
library(raster)
library(ggplot2)
library(hrbrthemes)
library(cowplot)
```

# Read data

Variables per plot:

Lauki1: 7594
Laukiz2: 7594
Oiartzun: 7841
Luiando: 7471

Why is there such a huge difference in the number of indices without NAs for the plots?

```{r}
data <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
coords <- readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
```

# Merge into one dataset

Leaving out Luiando (see README for reason).

```{r}
data <- as_tibble(rbindlist(data, fill = TRUE)) # combine all plots
data <- Filter(function(x) !any(is.na(x)), data) # filter out columsn containing NA
coords <- as_tibble(rbindlist(coords))
data %<>%
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(559, 382, 301, 499))))
```

# Mean defoliation per plot

```{r}
data %>%
  group_by(plot) %>%
  summarise(mean(deftot))
```

```{r}
data %>%
  group_by(plot) %>%
  ggplot(aes(y = deftot, x = plot)) +
  geom_boxplot() +
  theme_ipsum_rc()
```

# Point density

In Meters.

```{r}
plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")

dist <- map(plots, ~ {
  coords_laukiz1 <- coords %>%
    mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"), c(559, 382, 301, 499)))) %>%
    filter(plot == .x) %>%
    dplyr::select(-plot)

  laukiz1_points <- SpatialPoints(coords = coords_laukiz1, proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))

  distance <- mean(as.dist(pointDistance(laukiz1_points, allpairs = TRUE, lonlat = FALSE)))
})
set_names(dist, plots)
```

# Scatterplots

Visualize RMSE vs plot characteristics

```{r}
df <- tibble(
  plot = factor(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")),
  rmse = c(89.89, 30.37, 76.02, 106.65),
  point_dens = c(34.34, 62.51, 33.76, 35.01),
  mean_defol = c(53.22, 17.73, 68.35, 44.63)
) 
```

## RMSE vs point dens

```{r}
rmse_dens <- ggplot(df, aes(x = point_dens, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean point density (m)", y = "RMSE") +
  theme_ipsum_rc() +
  theme(legend.position = "none") +
  theme(plot.margin = unit(c(6,0,6,0), "pt"))
```

## RMSE vs mean defol

```{r}
rmse_defol <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean defoliation (%)", y = "RMSE") +
  theme_ipsum_rc() +
  theme(legend.position = "none") +
  theme(plot.margin = unit(c(6,0,6,0), "pt"))
```

## Plot

### Legend

```{r}
legend <- ggplot(df, aes(x = mean_defol, y = rmse)) +
  geom_point(aes(color = plot), shape = 19) +
  labs(x = "Mean defoliation (%)", y = "RMSE", color = "Plot") +
  theme_ipsum_rc()
legend <- get_legend(legend)
```

### Grid

```{r}
plot_grid(rmse_dens, legend, rmse_defol, ncol = 3, rel_widths = c(1, .3, 1))
```