### 2018-07-11 02_scripts/022_preprocessing.Rmd

---
title: "02-preprocessing"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r}
library(purrr)
library(tibble)
library(glue)
library(dplyr)
library(tidyr)
library(magrittr)
library(sf)
library(pbmcapply)
library(data.table)
```

# Read points with indices

Read the clean points with all extracted indices for each plot.
Indices are extracted in script `01_hyperspectral/06_extract_indices_*`.

```{r}
data = map(c("laukiz1", "laukiz2", "luiando", "oiartzun"), ~ readRDS(glue("/data/patrick/mod/tree-per-tree/2016/{.x}/all-indices-{.x}_bf2.rda")))
```

# Remove columns with NA values and "ID" columns

Ref: https://stackoverflow.com/questions/41343900/remove-row-columns-with-any-all-nan-values/41343980#41343980

luiando has some NA obs that need to be removed, otherwise the whole column will be removed by `Filter()`

```{r}
data[[3]] %<>% 
  na.omit()

data %<>% 
  map(~ dplyr::select(.x, -contains("_ID"))) %>% 
  map(~ as.data.table(.x)) %>% 
  #map(~ na.omit(.x)) %>% # some obs in luiando are NA after extraction
  map(~ dplyr::select(.x, -tree.number, -decoloration, -canker, -diameter, -height)) %>% 
  map(~ Filter(function(x) !any(is.na(x)), .x))
```

# Extract coordinates

The scaling of variables won't work with the `geom` column attached.
So we need to safe the coordinates, remove them and attach them afterwards again.
Coordinates are needed for the spatial partitioning during modeling.

```{r}
coords = map(data, ~ st_as_sf(.x, crs = 32630)) %>% 
  map(~ as.data.frame(st_coordinates(.x)))
coords = saveRDS(coords, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2-coords.rda")
data %<>% 
  map(~ st_as_sf(.x, crs = 32630)) %>% 
  map(~ st_set_geometry(.x, NULL)) 
```

# Standardize variables

```{r}
# data %<>% 
#   pbmclapply(function(x) mutate_at(x, .funs = scale, 
#                    .vars = vars(-deftot)),
#              mc.cores = 4
#       )
# saveRDS(data, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")

# data.table so much faster
data %<>% 
  pbmclapply(function(x) {
    cols = names(x)[2:length(names(x))]
    x[, (cols) := lapply(.SD, scale), .SDcols=cols]
  }, mc.cores = 4
      )
saveRDS(data, "/data/patrick/mod/hyperspectral/data_clean_with_indices/data_clean_standardized_bf2.rda")
```