### 2018-07-11 02_scripts/06_important_bands.Rmd

---
title: "06-Important bands"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(ggpubr)
library(dplyr)
library(stringr)
library(purrr)
library(glue)
library(cowplot)
```

Read data

```{r}
ls = map(c("laukiz1", "laukiz2", "luiando", "oiartzun", "all_plots"), 
         ~ readRDS(glue("/data/patrick/mod/hyperspectral/top-inds/top_inds_rr_{name}.rda", name = .x)))
```

#### Important spectral regions

Remove the "bf2" information and filter out NRI only.

```{r}
ls_inds = map(ls, ~ 
                .x %<>% 
                mutate(index = str_sub(index, 9, 17)) %>% 
                filter(str_detect(index, "b"))
)
```

Bring all plots on the same scale (0 -1)

```{r}
ls_inds %<>% 
  map(~ .x %<>% 
        mutate(coef = coef / max(coef)))
```

Extract all band numbers.

```{r}
ls_all_bands_rr =
  map(ls_inds, ~ .x %<>% 
        dplyr::select(index) %>% 
        str_match_all("[0-9]+") %>% 
        flatten() %>% 
        as.numeric())
```

Ok, bands 1-4 did not make it into the data (NA problems).

```{r}
map(ls_all_bands_rr, ~ .x %>% as.factor() %>% levels(.))
```

Create a tibble with the coefficients and the band numbers.

```{r}
ls_tibble_most_imp_all = map2(ls_inds, ls_all_bands_rr, ~
                            tibble(band = .y, weight = rep(.x$coef, times = 1, each = 2))
)
```

Create a tibble with the coefficients and the band numbers.

```{r}
ls_tibble_most_imp_sum = map2(ls_inds, ls_all_bands_rr, ~
                            tibble(band = .y, weight = rep(.x$coef, times = 1, each = 2)) %>% 
                            group_by(band) %>% 
                            summarise(score = sum(weight)) %>% 
                            mutate(band = band) %>% 
                            arrange(desc(score))
)
```

Classify the bands into spectral regions

1 - 60: visible (404 - 680)
61 - 71: red edge (680 - 730)
72 - 126: NIR (730 - 996)

```{r}
ls_tibble_most_imp_sum %<>% 
  map(~ .x %<>% 
        mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 nm - 680 nm)", 
                                           (band >= 61 & band <= 71) ~ "Red Edge (680 nm - 730 nm)",
                                           band >= 72 ~ "NIR (730 nm  - 996 nm)"))) %>% 
        arrange(desc(score))
  ) 
```

```{r}
ls_barplot = map2(ls_tibble_most_imp_sum, c("Laukiz 1", 
                                           "Laukiz 2", 
                                           "Luiando", 
                                           "Oiartzun", 
                                           "All Plots"), ~ 
                       ggbarplot(.x, x = "band", y = "score",                              # Color by groups
                                 add = "segments",                             # Add segments from y = 0 to dots
                                 fill = "group",
                                 color = "group",
                                 palette = "nejm",
                                 group = "band",
                                 title = .y,                              # Rotate vertically
                                 dot.size = 0.5,    
                                 width = 0.5,
                                 ylim = c(0, 50),# Large dot size
                                 ggtheme = theme_pubr(base_size = 4, margin = TRUE,
                                                      legend = "none")) + 
                       labs(y = "Score", color = "Spectral region") + 
                       scale_x_discrete(breaks = seq(0, 120, 20),
                                          labels = seq(0, 120, 20))
)
```

```{r}
legend_bar = ggbarplot(ls_tibble_most_imp_sum[[1]], x = "band", y = "score",                              # Color by groups
                                 add = "segments",                             # Add segments from y = 0 to dots
                                 fill = "group",
                                 palette = "nejm",
                                 group = "band",                            # Rotate vertically                    
                                 ggtheme = theme_pubr(base_size = 5, margin = TRUE,
                                                      legend = "right")) + 
                       labs(y = "Score", fill = "Spectral region")
```

## Barplot

```{r}
legend_plot_bar <- get_legend(legend_bar)
ls_barplot[[6]] = legend_plot_bar
plot_gridded_bar = plot_grid(plotlist = ls_barplot, ncol = 3)
save_plot(plot_gridded_bar,
          filename = here::here("03_figures/results/important-bands_barplot.png"))
```

## Boxplot

Classify the bands into spectral regions

1 - 60: visible (404 - 680)
61 - 71: red edge (680 - 730)
72 - 126: NIR (730 - 996)

```{r}
ls_tibble_most_imp_all %<>% 
  map(~ .x %<>% 
        mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 nm - 680 nm)", 
                                           (band >= 61 & band <= 71) ~ "Red Edge (680 nm - 730 nm)",
                                           band >= 72 ~ "NIR (730 nm  - 996 nm)"))) %>% 
        arrange(desc(weight))
  ) 
```

```{r}
ls_boxplot = map2(ls_tibble_most_imp_all, c("Laukiz 1", 
                                           "Laukiz 2", 
                                           "Luiando", 
                                           "Oiartzun", 
                                           "All Plots"), ~ 
                       ggboxplot(.x, x = "band", y = "weight",                              # Color by groups
                                 #add = "segments",                             # Add segments from y = 0 to dots
                                 #fill = "group",
                                 color = "group",
                                 palette = "nejm",
                                 group = "band",
                                 title = .y,                              # Rotate vertically
                                 width = 0.4,
                                 size = 0.1,
                                 outlier.size = 0.01,
                                 #size = 0.1, # Large dot size
                                 ggtheme = theme_pubr(base_size = 4, margin = TRUE,
                                                      legend = "none")) + 
                       labs(y = "Score", color = "Spectral region") + 
                       scale_x_discrete(breaks = seq(5, 125, 20),
                                          labels = seq(5, 125, 20))
)
```

```{r}
legend_box = ggboxplot(ls_tibble_most_imp_all[[1]], x = "band", y = "weight",                              # Color by groups
                                 #add = "segments",                             # Add segments from y = 0 to dots
                                 color = "group",
                                 palette = "nejm",
                                 group = "band",                            # Rotate vertically                    
                                 ggtheme = theme_pubr(base_size = 5, margin = TRUE,
                                                      legend = "right")) + 
                       labs(y = "Score", color = "Spectral region")
```

```{r}
legend_plot_box <- get_legend(legend_box)
ls_boxplot[[6]] = legend_plot_box
plot_gridded_box = plot_grid(plotlist = ls_boxplot, ncol = 3)
save_plot(plot_gridded_box, base_width = 12,
          filename = here::here("03_figures/results/important-bands_boxplot.png"))
```

## Top 50 indices only

Works only if the data arriving here was subsetted before, i.e. contains only the top XX bands.
Subsetting happens in the respective plot files, e.g. 03-modeling-laukiz1.Rmd.

Classify the bands into spectral regions

1 - 60: visible (404 - 680)
61 - 71: red edge (680 - 730)
72 - 126: NIR (730 - 996)

```{r}
ls_tibble_most_imp %<>% 
  map(~ .x %<>% 
        mutate(group = as.factor(case_when(band <= 60 ~ "Visible (404 nm - 680 nm)", 
                                           (band >= 61 & band <= 71) ~ "Red Edge (680 nm - 730 nm)",
                                           band >= 72 ~ "NIR (730 nm  - 996 nm)"))) %>% 
        arrange(desc(score)) %>% 
        mutate(band = as.factor(band))
  )
```

```{r}
ls_dotplots = map2(ls_tibble_most_imp, c("Laukiz 1 (bands = 56)", 
                                         "Laukiz 2 (bands = 41)", 
                                         "Luiando (bands = 61)", 
                                         "Oiartzun (bands = 79)", 
                                         "All Plots (bands = 76)"), ~ 
                     ggdotchart(.x, x = "band", y = "score",                              # Color by groups
                                sorting = "descending",                       # Sort value in descending order
                                add = "segments",                             # Add segments from y = 0 to dots
                                color = "group",
                                palette = "nejm",
                                position = position_dodge(1),
                                group = "band",
                                title = .y,
                                rotate = FALSE,                                # Rotate vertically
                                dot.size = 0.5,                                 # Large dot size
                                ggtheme = theme_pubr(base_size = 4, margin = TRUE,
                                                     legend = "none")) + 
                     labs(y = "Score", color = "Spectral region")
)

# adjust colors for Laukiz2 because "Red Edge" level is missing and hence the colors are different
ls_dotplots[[2]] = ls_dotplots[[2]] + 
  scale_color_manual(values = c("#619CFF", "#F8766D"))
```

```{r}
legend = ggdotchart(ls_tibble_most_imp[[1]], x = "band", y = "score",                              # Color by groups
                    sorting = "descending",                       # Sort value in descending order
                    add = "segments",                             # Add segments from y = 0 to dots
                    color = "group",
                    palette = "nejm",
                    group = "score",
                    rotate = TRUE,                                # Rotate vertically
                    dot.size = 0.5,                                 # Large dot size
                    #label = round(tibble_most_imp$sum),                        # Add mpg values as dot labels
                    ggtheme = theme_pubr(base_size = 3, margin = FALSE,
                                         legend = "right")) + 
  labs(y = "Score", color = "Spectral region")
```

```{r}
legend_plot <- get_legend(legend)
ls_dotplots[[6]] = legend_plot
plot_gridded = plot_grid(plotlist = ls_dotplots, ncol = 3)
save_plot(plot_gridded,
          filename = here::here("03_figures/results/important-bands.png"))
```

## Boxplot

```{r}
as_tibble(data.table::rbindlist(ls_tibble_most_imp)) %>% 
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun", "All Plots"), 
                           c(56, 41, 61, 79, 76)))) %>% 
  mutate(band = as.numeric(as.character(band))) %>% 
  group_by(plot) %>%
  ggboxplot(
    x = "plot", y = "band", color = "plot",
    add = "jitter")
```

```{r}
as_tibble(data.table::rbindlist(ls_tibble_most_imp)) %>% 
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun", "All Plots"), 
                           c(56, 41, 61, 79, 76)))) %>% 
  ggdotchart(x = "band", y = "score",                              # Color by groups
             sorting = "descending",                       # Sort value in descending order
             add = "segments",                             # Add segments from y = 0 to dots
             color = "group",
             palette = "nejm",
             group = "band",
             facet.by = "plot",
             rotate = F,                                # Rotate vertically
             dot.size = 0.5,                                 # Large dot size
             #label = round(tibble_most_imp$sum),                        # Add mpg values as dot labels
             ggtheme = theme_pubr(base_size = 4, margin = FALSE,
                                  legend = "right")) + 
  labs(y = "Score", color = "Spectral region")
```

## Histogram

```{r}
as_tibble(data.table::rbindlist(ls_tibble_most_imp)) %>% 
  mutate(plot = factor(rep(c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun", "All Plots"), 
                           c(56, 41, 61, 79, 76)))) %>% 
  mutate(band = as.numeric(as.character(band))) %>% 
  group_by(plot) %>%
  ggbarplot(
    x = "band", y = "score", color = "plot")
```

# PNG to PDF conversion

Converts all pngs in the specified folder to pdfs. 
`&&` makes the "convert" call conditional on `cd`.
`exec` reduces memory waste of shell.
Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.

```{r png to pdf, echo = FALSE, eval = FALSE}
# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
#        parallel convert '{}' '{.}.pdf'")

system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

# copy to presentation folder
# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
#           overwrite = TRUE)
```

# Session Info

```{r}
sessionInfo()
```