### 2018-07-31 02_scripts/08_prediction.Rmd
scripts

---
title: "08-prediction"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(purrr)
library(raster)
library(furrr)
library(sf)
library(pbmcapply)
```

# Load model

```{r}
mod = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned.rda")
```

# Load vegetation indices

```{r}
files = list.files("/data/patrick/mod/hyperspectral/all-veg-indices", 
                   pattern = ".grd", full.names = TRUE)

veg_inds = map(files, ~ brick(.))
```

## Format veg inds to sf objects

So that we have a data.frame with coordinates.

```{r}
plan(multicore)
veg_inds %<>% 
  future_map(~ as(., "SpatialPixelsDataFrame")) %>% 
  future_map(~ st_as_sf(.))
```

# Load NRI

```{r}
files_nri = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
                   pattern = ".grd", full.names = TRUE)

nri = map(files_nri, ~ brick(.))
```

## Format NRI to sf objects

So that we have a data.frame with coordinates.

```{r}
nri %<>% 
  pbmclapply(function(x) as(x, "SpatialPixelsDataFrame"), mc.cores = 2) %>% 
  pbmclapply(function(x) st_as_sf(x), mc.cores = 2)
```

# Name list elements

```{r}
names = map_chr(files, ~ tools::file_path_sans_ext(basename(.)))

nri %<>% 
  set_names(names)

veg_inds %<>% 
  set_names(names)
```

# Combine NRI and veg inds