### 2018-08-07 02_scripts/08_prediction.Rmd
update scripts

---
title: "08-prediction"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(purrr)
library(raster)
library(furrr)
library(sf)
library(pbmcapply)
library(magrittr)
library(dplyr)
library(data.table)
library(rasterVis)
library(viridis)
library(ggpubr)
```

# Load model

```{r}
mod = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned_7vars.rda")
```

# Load vegetation indices

```{r}
files = list.files("/data/patrick/mod/hyperspectral/all-veg-indices", 
                   pattern = ".grd", full.names = TRUE)

veg_inds = map(files, ~ brick(.))
```

## Format veg inds to sf objects

So that we have a data.frame with coordinates.

```{r}
plan(multicore)
veg_inds %<>% 
  future_map(~ as(., "SpatialPixelsDataFrame")) %>% 
  future_map(~ st_as_sf(.))
```

Prefix all vars with "bf2" so that they match with the names of the fitted model

```{r}
veg_inds %<>% 
  map(~ select_all(.x, .funs = funs(paste0("bf2_", .))))
```

Use only the indices used for training

```{r}
veg_inds %<>% 
  map(~ dplyr::select(.x, bf2_EVI, bf2_GDVI_4, bf2_GDVI_3, bf2_GDVI_2, 
                      bf2_mNDVI, bf2_mSR, bf2_D1))
```

# Load NRI

We only use the 7 most important variables. No NRI among them.

```{r}
# files_nri = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
#                    pattern = ".grd", full.names = TRUE)
# 
# nri = map(files_nri, ~ brick(.))
```

## Format NRI to sf objects

So that we have a data.frame with coordinates.

```{r}
# if (file.exists("/data/patrick/mod/hyperspectral/prediction/nri_basque.rda")) {
#   nri = readRDS("/data/patrick/mod/hyperspectral/prediction/nri_basque.rda")
# } else {
#   nri %<>% 
#     pbmclapply(function(x) as(x, "SpatialPixelsDataFrame"), mc.cores = 2) %>% 
#     pbmclapply(function(x) st_as_sf(x), mc.cores = 2)
# }
```

# Name list elements

```{r}
names = map_chr(files, ~ tools::file_path_sans_ext(basename(.)))

# nri %<>% 
#   set_names(names)

veg_inds %<>% 
  set_names(names)
```

<!-- # Combine NRI and veg inds -->

<!-- ```{r} -->
<!-- if (file.exists("/data/patrick/mod/hyperspectral/prediction/prediction_df.rda")) { -->
<!--   inds_combined = readRDS("/data/patrick/mod/hyperspectral/prediction/prediction_df.rda") -->
<!-- } else { -->
<!--   #options(future.globals.maxSize = 55000 * 1024^2) -->
<!--   #plan(multicore, workers = 2) -->
<!--   inds_combined = map2(veg_inds, nri, ~ st_join(.x, .y)) -->
<!--   saveRDS(inds_combined, "/data/patrick/mod/hyperspectral/prediction/prediction_df.rda") -->
<!-- } -->
<!-- ``` -->

<!-- Prefix all vars with "bf2" so that they match with the names of the fitted model -->

<!-- ```{r} -->
<!-- inds_combined %<>%  -->
<!--   map(~ select_all(.x, .funs = funs(paste0("bf2_", .)))) -->
<!-- ``` -->

<!-- Use only the indices used for training -->

<!-- ```{r} -->
<!-- indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda") -->

<!-- inds_combined %<>%  -->
<!--   map(~ dplyr::select(.x, indices_shared)) -->
<!-- ``` -->

Extract coordinates of all plots

```{r}
pred_df_x = map(veg_inds, ~ as.integer(st_coordinates(.x)[, 1]))
pred_df_y = map(veg_inds, ~ as.integer(st_coordinates(.x)[, 2]))

coords2 = map2(pred_df_x, pred_df_y, ~ data.frame(.x, .y))
coords2 = map(coords2, ~ dplyr::rename(.x, x = .x, y = .y))
```

# Predict

Make prediction data.frame

```{r}
veg_inds %<>% 
  map(~ st_set_geometry(.x, NULL))

saveRDS(veg_inds, "/data/patrick/mod/modeling/paper1_2018/prediction_df.rda")
```

# Standardize variables

```{r}
veg_inds %<>% 
  map(~ as.data.table(.x)) %>% 
  pbmclapply(function(x) {
    cols = names(x)[2:length(names(x))]
    x[, (cols) := lapply(.SD, scale), .SDcols=cols]
  }, mc.cores = 28
      ) %>% 
  map(~ as(.x, "data.frame"))
```

```{r}
#plan(multicore, workers = 28)

# correct column order of features: https://github.com/dmlc/xgboost/issues/1809
veg_inds %<>% 
  map(~ .x[mod$learner.model$feature_names])

pred = map(veg_inds, ~ predict(mod$learner.model, newdata = as.matrix(.x)))
```

# Combine prediction results with coords

```{r}
all = map2(coords2, pred, ~ cbind(.x, .y)) %>% 
  map(~ rename(.x, defoliation = .y))
```

# Transfer to SPDF and export raster

```{r}
all_spdf = map(all, ~ SpatialPixelsDataFrame(points = .x[, c("x", "y")], 
                                             data = as.data.frame(.x[, "defoliation"]), 
                                             proj4string = CRS("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"))) %>% 
  map(~ set_names(.x, "defoliation"))

# create raster objects
raster_list <- map(all_spdf, ~ raster(.x))

# write out names as there are some probs doing it dynamically in walk
names = map_chr(files, ~ tools::file_path_sans_ext(basename(.x)))

# WHY DOES this work??
iwalk(raster_list, ~ writeRaster(.x, glue::glue("/data/patrick/mod/hyperspectral/prediction/defol-pred-{.y}.tif"),
                          overwrite = TRUE)
)
```

# Plot demonstration plots

Plot title only

```{r}
# lattice.options(
#   layout.heights=list(bottom.padding=list(x=0), top.padding=list(x=0)),
#   layout.widths=list(left.padding=list(x=0), right.padding=list(x=0))
# )

raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

name = c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")

maps = imap(raster_list_wgs84[c(15, 16, 18, 19)], ~ levelplot( 
  .x[["defoliation"]], 
  #main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
  main = glue::glue("{tools::toTitleCase(.y)}"), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0), 
  at = seq(30, 80, length.out = 100),
  colorkey = list(
    width = 0.75,                       
    space = 'right',  
    labels = list(#at = c(0, 25, 50, 75, 100),
                font = 4),      
    axis.line = list(col = 'black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis)    
)
```

```{r}
png("../03_figures/results/predictions_demon_plots.png", units = "in", width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps, nrow = 2, ncol = 2)
dev.off()

```

Descriptive title

```{r}
# lattice.options(
#   layout.heights=list(bottom.padding=list(x=0), top.padding=list(x=0)),
#   layout.widths=list(left.padding=list(x=0), right.padding=list(x=0))
# )

raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

name = c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")

maps = imap(raster_list_wgs84[c(15, 16, 18, 19)], ~ levelplot( 
  .x[["defoliation"]], 
  main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
  #main = glue::glue("{tools::toTitleCase(.y)}"), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0), 
  at = seq(30, 80, length.out = 100),
  colorkey = list(
    width = 0.75,                       
    space = 'right',  
    labels = list(#at = c(0, 25, 50, 75, 100),
                font = 4),      
    axis.line = list(col = 'black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis)    
)
```

```{r}
png("../03_figures/results/predictions_demon_plots_descriptive_title.png", units = "in", width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps, nrow = 2, ncol = 2)
dev.off()

```

## Histograms

```{r}
df_all= map(all_spdf[c(15, 16, 18, 19)], ~ as(.x, "data.frame"))

hists_all = imap(df_all, ~ gghistogram(.x, x = "defoliation", 
                                       title = tools::toTitleCase(.y)))

hists_grid = patchwork::wrap_plots(hists_all)

cowplot::save_plot(hists_grid, filename = 
                   "../03_figures/results/predictions_demon_plots_hists.png", 
                   ncol = 2, nrow = 2,
                   )
```

# PNG to PDF conversion

Converts all pngs in the specified folder to pdfs. 
`&&` makes the "convert" call conditional on `cd`.
`exec` reduces memory waste of shell.
Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.

```{r png to pdf, echo = FALSE, eval = FALSE}
# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
#        parallel convert '{}' '{.}.pdf'")

system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

# copy to presentation folder
# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
#           overwrite = TRUE)
```

# Session Info

```{r}
sessionInfo()
```