### 2018-11-28 02_scripts/version-1/08_prediction.Rmd

---
title: "08-prediction"
output: 
  bookdown::html_document2:
    toc: true
    toc_float: true
---

# Load packages

```{r, results='hide'}
library(purrr)
library(raster)
library(furrr)
library(sf)
library(pbmcapply)
library(magrittr)
library(dplyr)
library(data.table)
library(rasterVis)
library(viridis)
library(ggpubr)
```

# Load model

```{r}
mod = readRDS("/data/patrick/mod/hyperspectral/prediction/xgboost_trained_tuned_7vars.rda")
```

# Load vegetation indices

```{r}
files = list.files("/data/patrick/mod/hyperspectral/all-veg-indices", 
                   pattern = ".grd", full.names = TRUE)

veg_inds = map(files, ~ brick(.))
```

## Format veg inds to sf objects

So that we have a data.frame with coordinates.

```{r}
plan(multicore)
veg_inds %<>% 
  future_map(~ as(., "SpatialPixelsDataFrame")) %>% 
  future_map(~ st_as_sf(.))
```

Prefix all vars with "bf2" so that they match with the names of the fitted model

```{r}
veg_inds %<>% 
  map(~ select_all(.x, .funs = funs(paste0("bf2_", .))))
```

Use only the indices used for training

```{r}
veg_inds %<>% 
  map(~ dplyr::select(.x, bf2_EVI, bf2_GDVI_4, bf2_GDVI_3, bf2_GDVI_2, 
                      bf2_mNDVI, bf2_mSR, bf2_D1))
```

# Load NRI

We only use the 7 most important variables. No NRI among them.

```{r}
# files_nri = list.files("/data/patrick/mod/hyperspectral/narrow-band-indices/", 
#                    pattern = ".grd", full.names = TRUE)
# 
# nri = map(files_nri, ~ brick(.))
```

## Format NRI to sf objects

So that we have a data.frame with coordinates.

```{r}
# if (file.exists("/data/patrick/mod/hyperspectral/prediction/nri_basque.rda")) {
#   nri = readRDS("/data/patrick/mod/hyperspectral/prediction/nri_basque.rda")
# } else {
#   nri %<>% 
#     pbmclapply(function(x) as(x, "SpatialPixelsDataFrame"), mc.cores = 2) %>% 
#     pbmclapply(function(x) st_as_sf(x), mc.cores = 2)
# }
```

# Name list elements

```{r}
names = map_chr(files, ~ tools::file_path_sans_ext(basename(.)))

# nri %<>% 
#   set_names(names)

veg_inds %<>% 
  set_names(names)
```

<!-- # Combine NRI and veg inds -->

<!-- ```{r} -->
<!-- if (file.exists("/data/patrick/mod/hyperspectral/prediction/prediction_df.rda")) { -->
<!--   inds_combined = readRDS("/data/patrick/mod/hyperspectral/prediction/prediction_df.rda") -->
<!-- } else { -->
<!--   #options(future.globals.maxSize = 55000 * 1024^2) -->
<!--   #plan(multicore, workers = 2) -->
<!--   inds_combined = map2(veg_inds, nri, ~ st_join(.x, .y)) -->
<!--   saveRDS(inds_combined, "/data/patrick/mod/hyperspectral/prediction/prediction_df.rda") -->
<!-- } -->
<!-- ``` -->

<!-- Prefix all vars with "bf2" so that they match with the names of the fitted model -->

<!-- ```{r} -->
<!-- inds_combined %<>%  -->
<!--   map(~ select_all(.x, .funs = funs(paste0("bf2_", .)))) -->
<!-- ``` -->

<!-- Use only the indices used for training -->

<!-- ```{r} -->
<!-- indices_shared = readRDS("/data/patrick/mod/hyperspectral/data_clean_with_indices/indices-shared-by-all-plots-bf2.rda") -->

<!-- inds_combined %<>%  -->
<!--   map(~ dplyr::select(.x, indices_shared)) -->
<!-- ``` -->

Extract coordinates of all plots

```{r}
pred_df_x = map(veg_inds, ~ as.integer(st_coordinates(.x)[, 1]))
pred_df_y = map(veg_inds, ~ as.integer(st_coordinates(.x)[, 2]))

coords2 = map2(pred_df_x, pred_df_y, ~ data.frame(.x, .y))
coords2 = map(coords2, ~ dplyr::rename(.x, x = .x, y = .y))
```

# Predict

Make prediction data.frame

```{r}
veg_inds %<>% 
  map(~ st_set_geometry(.x, NULL))

saveRDS(veg_inds, "/data/patrick/mod/modeling/paper1_2018/prediction_df.rda")
```

# Standardize variables

```{r}
veg_inds %<>% 
  map(~ as.data.table(.x)) %>% 
  pbmclapply(function(x) {
    cols = names(x)[2:length(names(x))]
    x[, (cols) := lapply(.SD, scale), .SDcols=cols]
  }, mc.cores = 28
      ) %>% 
  map(~ as(.x, "data.frame"))
```

```{r}
#plan(multicore, workers = 28)

# correct column order of features: https://github.com/dmlc/xgboost/issues/1809
veg_inds %<>% 
  map(~ .x[mod$learner.model$feature_names])

pred = map(veg_inds, ~ predict(mod$learner.model, newdata = as.matrix(.x)))
```

# Combine prediction results with coords

```{r}
all = map2(coords2, pred, ~ cbind(.x, .y)) %>% 
  map(~ rename(.x, defoliation = .y))
```

# Transfer to SPDF and export raster

```{r}
all_spdf = map(all, ~ SpatialPixelsDataFrame(points = .x[, c("x", "y")], 
                                             data = as.data.frame(.x[, "defoliation"]), 
                                             proj4string = CRS("+proj=utm +zone=30 +datum=WGS84 +units=m +no_defs"))) %>% 
  map(~ set_names(.x, "defoliation"))

# create raster objects
raster_list <- map(all_spdf, ~ raster(.x))

# write out names as there are some probs doing it dynamically in walk
names = map_chr(files, ~ tools::file_path_sans_ext(basename(.x)))

# WHY DOES this work??
iwalk(raster_list, ~ writeRaster(.x, glue::glue("/data/patrick/mod/hyperspectral/prediction/defol-pred-{.y}.tif"),
                          overwrite = TRUE)
)
```

# Plot demonstration plots

Plot title only

```{r}
raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

name = c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")

maps = imap(raster_list_wgs84[c(15, 16, 18, 19)], ~ levelplot( 
  .x[["defoliation"]], 
  #main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
  main = glue::glue("{tools::toTitleCase(.y)}"), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0), 
  at = seq(30, 80, length.out = 100),
  colorkey = list(
    width = 0.75,                       
    space = 'right',  
    labels = list(#at = c(0, 25, 50, 75, 100),
                font = 4),      
    axis.line = list(col = 'black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis)    
)
```

```{r}
png("../03_figures/results/predictions_demon_plots.png", units = "in", width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps, nrow = 2, ncol = 2)
dev.off()
```

Descriptive title

```{r}
raster_list_wgs84 = map(raster_list, ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"))

name = c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")

maps = imap(raster_list_wgs84[c(15, 16, 18, 19)], ~ levelplot( 
  .x[["defoliation"]], 
  main = glue::glue("{tools::toTitleCase(.y)}: Predicted defoliation of trees (in %)"), 
  #main = glue::glue("{tools::toTitleCase(.y)}"), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0), 
  at = seq(30, 80, length.out = 100),
  colorkey = list(
    width = 0.75,                       
    space = 'right',  
    labels = list(#at = c(0, 25, 50, 75, 100),
                font = 4),      
    axis.line = list(col = 'black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis)    
)
```

```{r}
png("../03_figures/results/predictions_demon_plots_descriptive_title.png", units = "in", width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps, nrow = 2, ncol = 2)
dev.off()

```

## Histograms

```{r}
df_all= map(all_spdf[c(15, 16, 18, 19)], ~ as(.x, "data.frame"))

hists_all = imap(df_all, ~ gghistogram(.x, x = "defoliation", 
                                       title = tools::toTitleCase(.y)))

hists_grid = patchwork::wrap_plots(hists_all)

cowplot::save_plot(hists_grid, filename = 
                   "../03_figures/results/predictions_demon_plots_hists.png", 
                   ncol = 2, nrow = 2,
                   )
```

# Plot all plots

Descriptive title

```{r}
raster_list = map(list.files("/data/patrick/mod/hyperspectral/prediction/", 
                             pattern = "defol-pred*", full.names = TRUE), 
                  ~ raster(.x)
                  )
raster_list_wgs84 = map(raster_list, 
                        ~ projectRaster(.x, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")) %>% 
  map(~ set_names(.x, "defoliation"))

name = c("Aguinaga", "Altube", "Azaceta", "Barazar", "busturi", "Caranca",
         "Douglas 1", "Douglas 2", "Douglas Albiztur", "Douglas Legutio",
         "Gipuzkoa", "Laricio Defol Olaeta", "Laricio Defol Urkiola",
         "Laricio Gatzaga", "Laukiz 1", "Laukiz 2", "Laukiz 3", "Luiando", 
         "Oiartzun", "Oxtandio", "Picea Abornicano", "Pinas Santa Koloma",
         "Pinast Arza", "Rad Altube", "Rad Menagaray", "Sequoia Legorreta", 
         "Silvest Nograro")

maps = imap(raster_list_wgs84, ~ levelplot( 
  .x[["defoliation"]], 
  main = glue::glue("{tools::toTitleCase(name[.y])}: Defoliation [%]"), 
  #main = glue::glue("{tools::toTitleCase(.y)}"), 
  margin = FALSE,                             
  between = list(x = 0.0, y = 0.0), 
  at = seq(30, 82, length.out = 100), # min/max from map(raster_list_wgs84, ~ summary(.x))
  colorkey = list(
    width = 0.75,                       
    space = 'right',  
    labels = list(#at = c(0, 25, 50, 75, 100),
                font = 4),      
    axis.line = list(col = 'black')
  ),
  par.strip.text = list(col = "white", cex = 1.2),    
  col.regions = viridis)    
)
```

```{r}
png("../03_figures/results/predictions_plots-1-4.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[1:4], nrow = 2, ncol = 2)
dev.off()

png("../03_figures/results/predictions_plots-5-8.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[5:8], nrow = 2, ncol = 2)
dev.off()

png("../03_figures/results/predictions_plots-9-12.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[9:12], nrow = 2, ncol = 2)
dev.off()

png("../03_figures/results/predictions_plots-12-16-.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[13:16], nrow = 2, ncol = 2)
dev.off()

png("../03_figures/results/predictions_plots-17-20.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[17:20], nrow = 2, ncol = 2)
dev.off()

png("../03_figures/results/predictions_plots-21-24.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[21:24], nrow = 2, ncol = 2)
dev.off()

png("../03_figures/results/predictions_plots-25-28.png", units = "in", 
    width = 13, height = 9.21, res = 300)
gridExtra::grid.arrange(grobs = maps[25:28], nrow = 2, ncol = 2)
dev.off()
```

# PNG to PDF conversion

Converts all pngs in the specified folder to pdfs. 
`&&` makes the "convert" call conditional on `cd`.
`exec` reduces memory waste of shell.
Uses "GNU parallel" library for parallel jobs and the `convert` function from the `ImageMagick` library.

```{r png to pdf, echo = FALSE, eval = FALSE}
# system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/data && exec ls -1 *.png |
#        parallel convert '{}' '{.}.pdf'")

system("cd /home/patrick/PhD/papers/03_hyperspectral/03_figures/results/ && exec ls -1 *.png | 
       parallel convert '{}' '{.}.pdf'")

# copy to presentation folder
# file_copy("/home/patrick/PhD/papers/01_model_comparison/04_figures/02_results/cv_boxplots_final.png",
#           "/home/patrick/PhD/presentations/paper1/figs/cv_boxplots_final.png",
#           overwrite = TRUE)
```

# Session Info

```{r}
sessionInfo()
```