### 2019-02-01 README.Rmd
restructure

---
output: github_document
---

[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/pat-s/paper_hyperspectral/commits/master)  
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-brightgreen.svg)](https://cran.r-project.org/)   
[![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)  
<!-- [![Travis-CI Build Status](https://travis-ci.org/pat-s/paper_hyperspectral.png?branch=master)](https://travis-ci.org/pat-s/paper_hyperspectral)  -->

### Compendium URL

### Authors

**Patrick Schratz** (patrick.schratz@gmail.com) [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0003-0748-6624)    
Jannes Muenchow [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-7834-4717)   
Bernd Bischl [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6002-6980)   
Alexander Brenning [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6640-679X)

### Contents

This repository contains the research compendium of our work on comparing algorithms across multiple feature sets and filtering methods (including ensemble filter methods).

- Using machine-learning algorithms to model defoliation of _Pinus Radiata_ trees.

- keywords

    - high-dimensional data handling
    - feature selection
    - machine-learning
    - ecology

- Compare filtering methods (ensemble filter methods) across various algorithms and datasets

- Predict defoliation to all available plots (24) and the whole Basque Country (at 200 m resolution)

### How to use

#### Read the code, access the data

See the [`analysis`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis) directory on GitHub for the source code that generated the figures and statistical results contained in the manuscript.
See the [`data`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis/data) directory for instructions how to access the raw data discussed in the manuscript.

#### Install the R package

[![Build Status](https://travis-ci.org/pat-s/paper_hyperspectral.svg?branch=master)](https://travis-ci.org/pat-s/paper_hyperspectral)

This repository is organized as an R package, providing functions and raw data to reproduce and extend the analysis reported in the publication. 
Note that this package has been written explicitly for this project and may not be suitable for more general use.

This project is setup with a [drake workflow](https://github.com/ropensci/drake), ensuring reproducibility.
Intermediate targets/objects will be stored in a hidden `.drake` directory.

The R library of this project is managed by [packrat](https://rstudio.github.io/packrat/).
This makes sure that the exact same package versions are used when recreating the project.
When calling `packrat::restore()`, all required packages will be installed with their specific version.

Please note that this project was built with R version 3.5.2 on a Debian 9 operating system.
The packrat packages from this project **are not compatible with R versions prior version 3.5.0.**
For reproducibility, it is recommended to replicate the analysis using the included Dockerfile.
Instructions can be found [ħere](https://github.com/pat-s/paper_hyperspectral#docker).
(In general, it should be possible to reproduce the analysis on any other operating system.)

To clone the project, a working installation of `git` is required.
Open a terminal in the directory of your choice and execute:

```sh
git clone git@github.com:pat-s/paper_hyperspectral.git
```

Then start R in this directory and run

```{r README-1, eval = FALSE}
packrat::restore()
source("analysis/drake.R")
make(plan, keep_going = TRUE, console_log_file = stdout()) 
# use more cores with make(plan, jobs = <number of cores>)
```

## Runtime

Predicted total runtime (based on all target runtimes ever built and stored in the internal `config` file)

```{r README-2 }
source("analysis/drake.R")
predict_runtime(config, from_scratch = TRUE)
```

Acceleration when parallelizing the `make()` call

```{r README-3, warning=FALSE, message=FALSE}
time <- c()
for (jobs in 1:10){
time[jobs] <- predict_runtime(
drake_config(),
jobs = jobs,
from_scratch = TRUE,
known_times = build_times(targets_only = TRUE)$elapsed
)
}

library(ggplot2)
ggplot(data.frame(time = time / 3600, jobs = ordered(1:10), group = 1)) +
geom_line(aes(x = jobs, y = time, group = group)) +
scale_y_continuous(breaks = 0:10 * 4, limits = c(0, 29)) +
ggpubr::theme_pubr() +
xlab("jobs argument of make()") +
ylab("Predicted runtime of make() (hours)")
```

### Licenses

Text: CC-BY-4.0 http://creativecommons.org/licenses/by/4.0/

Code: MIT http://opensource.org/licenses/MIT year: 2019, copyright holder: Patrick Schratz

Data: CC0 http://creativecommons.org/publicdomain/zero/1.0/

### Notes and resources

* The [issues tracker](https://github.com/pat-s/paper_hyperspectral/issues) is the place to report problems or ask questions

* See the repository [history](https://github.com/pat-s/paper_hyperspectral/commits/master) for a fine-grained view of progress and changes.

* The organisation of this compendium is based on the work of [Carl Boettiger](http://www.carlboettiger.info/) and [Ben Marwick](https://github.com/benmarwick).

## Literature

- What to do first (Feature selection or tuning): https://datascience.stackexchange.com/a/19778/24475
- Hastie, T., Friedman, J., & Tibshirani, R. (2001). The Elements of Statistical Learning. Springer New York. https://doi.org/10.1007/978-0-387-21606-5 (Kapitel 16)
- FSel algs overview: http://featureselection.asu.edu/algorithms.php
- De Jay, N., Papillon-Cavanagh, S., Olsen, C., El-Hachem, N., Bontempi, G., & Haibe-Kains, B. (2013). mRMRe: an R package for parallelized mRMR ensemble feature selection. Bioinformatics, 29(18), 2365–2368. https://doi.org/10/f48vxc
- Drotár, P., Gazda, J., & Smékal, Z. (2015). An experimental comparison of feature selection methods on two-class biomedical datasets. Computers in Biology and Medicine, 66, 1–10. https://doi.org/10/f72ttc
- Drotár, P., Šimoňák, S., Pietriková, E., Chovanec, M., Chovancová, E., Ádám, N., … Biňas, M. (2017). Comparison of Filter Techniques for Two-Step Feature Selection. COMPUTING AND INFORMATICS, 36(3), 597–617. https://doi.org/10/gbntp7
- Drotár, P., Gazda, M., & Gazda, J. (2017). Heterogeneous ensemble feature selection based on weighted Borda count. In 2017 9th International Conference on Information Technology and Electrical Engineering (ICITEE) (pp. 1–4). https://doi.org/10.1109/ICITEED.2017.8250495
- Nie, F., Huang, H., Cai, X., & Ding, C. H. (2010). Efficient and Robust Feature Selection via Joint \mathscrl2,1-Norms Minimization. In J. D. Lafferty, C. K. I. Williams, J. Shawe-Taylor, R. S. Zemel, & A. Culotta (Eds.), Advances in Neural Information Processing Systems 23 (pp. 1813–1821). Curran Associates, Inc. Retrieved from http://papers.nips.cc/paper/3988-efficient-and-robust-feature-selection-via-joint-l21-norms-minimization.pdf