### 2019-04-10 analysis/rmd/defoliation-maps.Rmd
add analysis/rmd

---
title: "Creation of prediction maps"
output: html_document
author: "Patrick Schratz, FSU Jena"
---

```{r setup}
knitr::opts_knit$set(base.dir = 'analysis/rmd/')
knitr::opts_chunk$set(fig.retina=3, fig.align = 'center',
                      out.width = "100%")
loadd(defoliation_raster_2017, defoliation_raster_2018,
      defoliation_df_2017, defoliation_df_2018)
```

# Defoliation evaluation {.tabset .tabset-fade}

## Maps

```{r defoliation-map-2017, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
create_prediction_map(list(defoliation_raster_2017), "xgboost")
```

```{r defoliation-map-2018, fig.height = 5.5, fig.width = 8.5, fig.path="../../analysis/figures/", dev = c("png")}
create_prediction_map(list(defoliation_raster_2018), "xgboost")
```

## Histograms

```{r}
ggplot(defoliation_df_2017, aes(x = defoliation)) +
  geom_histogram() +
  labs(x = "Defoliation (%)") +
  theme_pubr()
```

```{r}
ggplot(defoliation_df_2018, aes(x = defoliation)) +
  geom_histogram() +
  labs(x = "Defoliation (%)") +
  theme_pubr()
```