### 2019-04-12 analysis/rmd/timeline.Rmd
add timeline.Rmd

---
title: "Timeline"
output: html_document
author: "Patrick Schratz, FSU Jena"
---

```{r setup, echo = FALSE}
knitr::opts_knit$set(base.dir = 'analysis/rmd/')
knitr::opts_chunk$set(fig.retina=3, fig.align = 'center',
                      out.width = "100%")
```

```{r}
library(vistime)
library(tidyverse)
library(lubridate)

d = read.csv(stringsAsFactors = FALSE, text = "event,start,duration,group
structure,0,2,orga
polish repo/code,10,3,orga
select journal,9,1.2,orga
drake pipeline,0,3,coding
finish benchmarks,1,10,coding
figures and tables,10,4,coding
literature,5,3,writing
review #1,12,1.2,writing
optimize,13,1.2,writing
review #2,14,1.2,writing
optimize,15,1.2,writing
first draft,9,3.1,writing
submit,16,1,writing"
)

start_date = as_date(lubridate::ymd("2019-04-08", tz = "GMT"))
d$start = start_date + d$start * 7
d$end = d$start + d$duration * 7
vistime(d, linewidth = 20)
```