### 2019-06-28 analysis/filter-correlation.Rmd
export filter-correlation images

---
title: "Correlation of filter methods"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

source(here::here("code/99-packages.R"))
sourceDirectory("R")

# load drake objects
loadd(
  fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
  fv_nri_symuncert, fv_nri_relief,
  fv_nri_rank, fv_nri_cor,
  fv_nri_mrmr,
  fv_nri_cmim, fv_nri_var,

  fv_vi_car, fv_vi_info.gain, fv_vi_gain.ratio,
  fv_vi_symuncert, fv_vi_relief,
  fv_vi_rank, fv_vi_cor,
  fv_vi_mrmr,
  fv_vi_cmim, fv_vi_var,

  fv_hr_car, fv_hr_info.gain, fv_hr_gain.ratio,
  fv_hr_symuncert, fv_hr_relief,
  fv_hr_rank, fv_hr_cor,
  fv_hr_mrmr,
  fv_hr_cmim, fv_hr_var,

  fv_hr_info.gain5, fv_hr_info.gain10,
  fv_hr_info.gain15, fv_hr_info.gain20,
  fv_hr_info.gain25, fv_hr_info.gain30,

  fv_vi_info.gain5, fv_vi_info.gain10,
  fv_vi_info.gain15, fv_vi_info.gain20,
  fv_vi_info.gain25, fv_vi_info.gain30,

  fv_nri_info.gain5, fv_nri_info.gain10,
  fv_nri_info.gain15, fv_nri_info.gain20,
  fv_nri_info.gain25, fv_nri_info.gain30
)
library("rgdal")
```

# Correlation of filter methods {.tabset .tabset-fade}

Spearman's rank correlation is used because we compare rankings against each other.

## VI

### Filter methods amongst each other

```{r correlation-filter-vi, warning=FALSE, dev = c("png", "pdf")}
# combine filter values

fv_all <- dplyr::bind_rows(
  fv_vi_car$data,
  fv_vi_info.gain$data,
  fv_vi_gain.ratio$data,
  fv_vi_symuncert$data,
  fv_vi_relief$data,
  fv_vi_rank$data,
  fv_vi_cor$data,
  fv_vi_mrmr$data,
  fv_vi_cmim$data,
  fv_vi_var$data
) %>%
  dplyr::select(value, method) %>%
  mutate(id = rep(1:89, 10)) %>%
  tidyr::spread(method, value) %>%
  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
  na.omit() %>% # NA in cmim
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")
# p.mat = cor_pmat(corr)

ggcorrplot(corr,
           # p.mat = p.mat,
           hc.order = F,
           type = "lower",
           lab = TRUE
)
```

### Number of bins of `FSelectorRcpp::information.gain()`

```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
# recode levels so that they are unique
fv_all_list = map2(list(fv_vi_info.gain5$data, fv_vi_info.gain10$data,
  fv_vi_info.gain15$data, fv_vi_info.gain20$data,
  fv_vi_info.gain25$data, fv_vi_info.gain30$data),
  list(5, 10, 15, 20, 25, 30), ~
    dplyr::mutate(.x, method = recode(.x$method, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}"))))

fv_all <- dplyr::bind_rows(
  fv_all_list
)  %>%
  dplyr::select(value, method) %>%
  mutate(id = rep(1:89, 6)) %>%
  tidyr::spread(method, value) %>%
  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
  na.omit() %>% 
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")

ggcorrplot(corr,
           # p.mat = p.mat,
           hc.order = F,
           type = "lower",
           lab = TRUE
)
```

## NRI

```{r filter-correlation-3, warning=FALSE}
# combine filter values

fv_all <- dplyr::bind_rows(
  fv_nri_car$data,
  fv_nri_info.gain$data,
  fv_nri_gain.ratio$data,
  fv_nri_symuncert$data,
  fv_nri_relief$data,
  fv_nri_rank$data,
  fv_nri_cor$data,
  fv_nri_mrmr$data,
  fv_nri_cmim$data,
  fv_nri_var$data
) %>%
  dplyr::select(value, method) %>%
  mutate(id = rep(1:7381, 10)) %>%
  tidyr::spread(method, value) %>%
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")
p.mat <- cor_pmat(corr)

ggcorrplot(corr,
           # p.mat = p.mat,
           hc.order = FALSE,
           type = "lower",
           lab = TRUE
)
```

## HR

```{r filter-correlation-4, warning=FALSE}
# combine filter values

fv_all <- dplyr::bind_rows(
  fv_hr_car$data,
  fv_hr_info.gain$data,
  fv_hr_gain.ratio$data,
  fv_hr_symuncert$data,
  fv_hr_relief$data,
  fv_hr_rank$data,
  fv_hr_cor$data,
  fv_hr_mrmr$data,
  fv_hr_cmim$data,
  fv_hr_var$data
) %>%
  dplyr::select(value, method) %>%
  mutate(id = rep(1:122, 10)) %>%
  tidyr::spread(method, value) %>%
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")
p.mat <- cor_pmat(corr)

ggcorrplot(corr,
           # p.mat = p.mat,
           hc.order = F,
           type = "lower",
           lab = TRUE
)
```

```{r filter-correlation-5, echo = FALSE}
file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
          "code/98-paper/journal/")
file_move("docs/figure/filter-correlation.Rmd/correlation-filter-vi-1.pdf",
          "code/98-paper/journal/")
```