### 2020-01-15 analysis/filter-correlation.Rmd
wflow_publish(knitr_in("analysis/filter-correlation.Rmd"), view = FALSE, verbose = TRUE)

---
title: "Correlation of filter methods"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

# source(here::here("code/99-packages.R"))
library("dplyr")
library("tidyr")
library("ggcorrplot")
library("purrr")
library("glue")
library("fs")
R.utils::sourceDirectory("R")

# load drake objects
drake::loadd(
  fv_nri_car, fv_nri_info.gain, fv_nri_gain.ratio,
  fv_nri_symuncert, fv_nri_relief,
  fv_nri_rank, fv_nri_cor,
  fv_nri_mrmr,
  fv_nri_cmim, fv_nri_var,

  fv_vi_car, fv_vi_info.gain, fv_vi_gain.ratio,
  fv_vi_symuncert, fv_vi_relief,
  fv_vi_rank, fv_vi_cor,
  fv_vi_mrmr,
  fv_vi_cmim, fv_vi_var,

  fv_hr_car, fv_hr_info.gain, fv_hr_gain.ratio,
  fv_hr_symuncert, fv_hr_relief,
  fv_hr_rank, fv_hr_cor,
  fv_hr_mrmr,
  fv_hr_cmim, fv_hr_var,

  fv_hr_info.gain5, fv_hr_info.gain10,
  fv_hr_info.gain15, fv_hr_info.gain20,
  fv_hr_info.gain25, fv_hr_info.gain30,

  fv_vi_info.gain5, fv_vi_info.gain10,
  fv_vi_info.gain15, fv_vi_info.gain20,
  fv_vi_info.gain25, fv_vi_info.gain30,

  fv_nri_info.gain5, fv_nri_info.gain10,
  fv_nri_info.gain15, fv_nri_info.gain20,
  fv_nri_info.gain25, fv_nri_info.gain30
)
```

# Correlation of filter methods {.tabset .tabset-fade}

Spearman's rank correlation is used because rankings are compared.

## VI

### Filter methods amongst each other

```{r filter-correlation-1, warning=FALSE}
# combine filter values

fv_all <- dplyr::bind_rows(
  fv_vi_car$data,
  fv_vi_info.gain$data,
  fv_vi_gain.ratio$data,
  fv_vi_symuncert$data,
  fv_vi_relief$data,
  fv_vi_rank$data,
  fv_vi_cor$data,
  fv_vi_mrmr$data,
  fv_vi_cmim$data,
  fv_vi_var$data
) %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:89, 10)) %>%
  tidyr::spread(filter, value) %>%
  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
  na.omit() %>% # NA in cmim
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")
# p.mat = cor_pmat(corr)

ggcorrplot(corr,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

### Number of bins of `FSelectorRcpp::information.gain()`

```{r correlation-nbins, warning=FALSE, dev = c("png", "pdf")}
# recode levels so that they are unique
fv_all_list <- map2(
  list(
    fv_vi_info.gain5$data, fv_vi_info.gain10$data,
    fv_vi_info.gain15$data, fv_vi_info.gain20$data,
    fv_vi_info.gain25$data, fv_vi_info.gain30$data
  ),
  list(5, 10, 15, 20, 25, 30), ~
  dplyr::mutate(.x, filter = recode(.x$filter, FSelectorRcpp_information.gain = glue("FSelectorRcpp_information.gain-{.y}")))
)

fv_all <- dplyr::bind_rows(
  fv_all_list
) %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:89, 6)) %>%
  tidyr::spread(filter, value) %>%
  mutate_if(is.numeric, list(~ na_if(., -Inf))) %>% # Inf in mrmr
  na.omit() %>%
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")

ggcorrplot(corr,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

## NRI

```{r correlation-filter-nri, dev = c("png", "pdf"), warning=FALSE}
# combine filter values

fv_all <- dplyr::bind_rows(
  fv_nri_car$data,
  fv_nri_info.gain$data,
  fv_nri_gain.ratio$data,
  fv_nri_symuncert$data,
  fv_nri_relief$data,
  fv_nri_rank$data,
  fv_nri_cor$data,
  fv_nri_mrmr$data,
  fv_nri_cmim$data,
  fv_nri_var$data
) %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:7381, 10)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")
p.mat <- cor_pmat(corr)

ggcorrplot(corr,
  # p.mat = p.mat,
  hc.order = FALSE,
  type = "lower",
  lab = TRUE
)
```

## HR

```{r filter-correlation-4, warning=FALSE}
# combine filter values

fv_all <- dplyr::bind_rows(
  fv_hr_car$data,
  fv_hr_info.gain$data,
  fv_hr_gain.ratio$data,
  fv_hr_symuncert$data,
  fv_hr_relief$data,
  fv_hr_rank$data,
  fv_hr_cor$data,
  fv_hr_mrmr$data,
  fv_hr_cmim$data,
  fv_hr_var$data
) %>%
  dplyr::select(value, filter) %>%
  mutate(id = rep(1:122, 10)) %>%
  tidyr::spread(filter, value) %>%
  dplyr::select(-id)

corr <- cor(fv_all, method = "spearman")
p.mat <- cor_pmat(corr)

ggcorrplot(corr,
  # p.mat = p.mat,
  hc.order = F,
  type = "lower",
  lab = TRUE
)
```

```{r filter-correlation-5, echo = FALSE}
system("sleep 5")
# file_move("docs/figure/filter-correlation.Rmd/correlation-nbins-1.pdf",
#           "code/98-paper/journal/")
# file_move("docs/figure/filter-correlation.Rmd/correlation-filter-nri-1.pdf",
#           "code/98-paper/journal/")
```