### 2020-03-08 analysis/inspect-glmnet-models.Rmd
wflow_publish(knitr_in("analysis/inspect-glmnet-models.Rmd"), view = FALSE, verbose = TRUE)

---
title: "Inspecting {glmnet} performances"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = TRUE
)

R.utils::sourceDirectory("R")
library("drake")
library("mlr")
library("glmnet")

# load drake objects
loadd(
  benchmark_models_new_penalized_buffer2,
  task_new_buffer2
)

```

Last update:

```{r}
date()
```

## Inspect fitted models during CV

Inspect Ridge regression on VI task in detail because the error is enourmus.

First extract the models.

```{r}
models = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["models"]]
```

Then look at the fold performances

```{r}
benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["measures.test"]][["rmse"]]
```

We see a high error on Fold 1 (= Laukiz 2).
The others are also quite high but not "out of bounds".

Because this models used the internal optimization of the lambda sequence (`cv.glmnet`), let's look at the value which was chosen for prediction (parameter `s` which defaults to `s="lambda.1se"`):

```{r}
purrr::map_dbl(models, ~ .x[["learner.model"]][["lambda.1se"]])
```

It seems that the `lambda.1se` value for Fold 1 is way smaller than for the other 3 folds.
However, all values seem to be quite high.

Let's look at the full lambda sequence

```{r}
purrr::map_int(models, ~ length(.x[["learner.model"]][["lambda"]]))
```

Interestingly, the lambda length of fold 1 is not 100 (default) but only 5.

## Train/predict manually for Fold 1 via {glmnet}

To inspect further, let's refit a {glmnet} model directly on the training data of Fold 1 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:

```{r}
train_inds_fold1 = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["pred"]][["instance"]][["train.inds"]][[1]]

obs_train_f1 = as.matrix(task_new_buffer2[[2]]$env$data[train_inds_fold1, getTaskFeatureNames(task_new_buffer2[[2]])])
target_f1 = getTaskTargets(task_new_buffer2[[2]])[train_inds_fold1]
```

Fit `cv.glmnet`

```{r}
set.seed(1)
modf1 = glmnet::cv.glmnet(obs_train_f1, target_f1, nfolds = 5)

modf1$lambda.1se
```

Ok, a value of `0.85` is **very** different to what happened during the CV (4.211054e+08).

Predict on Laukiz 2 now.

```{r}
pred_inds_fold1 = benchmark_models_new_penalized_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["pred"]][["instance"]][["test.inds"]][[1]]

obs_pred_f1 = as.matrix(task_new_buffer2[[2]]$env$data[pred_inds_fold1, getTaskFeatureNames(task_new_buffer2[[2]])])

pred = predict(modf1, newx = obs_pred_f1)
```

Calculate the error

```{r}
truth = task_new_buffer2[[2]]$env$data[pred_inds_fold1, "defoliation"]

mlr:::measureRMSE(truth, pred)
```

Ok, RMSE of 52. Still not great but at least not out of bounds.

## Train/predict manually for Fold 1 via {mlr}

Now let's fit the same subset (= Fold 1) via mlr

```{r}
lrn = makeLearner("regr.cvglmnet", nfolds = 5)
task_f1 = subsetTask(task_new_buffer2[[2]], train_inds_fold1)

set.seed(1)
mod = train(lrn, task_f1)
```

Check lambda sequence and `lambda.1se`:

```{r}
mod$learner.model$lambda
```

```{r}
mod$learner.model$lambda.1se
```

Check for equality between {mlr} and {glmnet} directly

```{r}
all.equal(modf1$lambda.1se, mod$learner.model$lambda.1se)
```

Ok, so {mlr} works correctly when fitting the model for a single fold. 
Hence, somethings goes wrong during CV?