### 2020-03-11 analysis/inspect-glmnet-models.Rmd
workflowr::wflow_publish("analysis/inspect-glmnet-models.Rmd")

---
title: "Inspecting {glmnet} performances"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = TRUE
)

R.utils::sourceDirectory("R")
library("drake")
library("mlr")
library("glmnet")
library("ggplot2")
library("magrittr")

options(crayon.enabled = TRUE, pillar.bold = TRUE, scipen = 999)
fansi::set_knit_hooks(knitr::knit_hooks)

# load drake objects
loadd(
  benchmark_models_new_penalized_mbo_buffer2,
  task_new_buffer2
)
```

Last update:

```{r, echo=FALSE, results='asis'}
date()
```

## General glmnet notes

- {glmnet} does its own internal 10-fold CV optimization when using `cv.glmnet()`.
  It iterates over `lambda` and chooses the most robust values for prediction via parameter `s` in `predict.glmnet()`. Supplying a custom lambda sequence does not make much sense since the internal heuristics are quite good (if one wants to use non-spatial optimization).
- To conduct a spatial optimization, one needs to use `glmnet()` directly. 
  This implementation does not do an internal optimization for `lambda`  and hence `s` can/needs to be tuned directly by the user. Because it is hard to come up with good tuning ranges in this case, one can fit a `cv.glmnet()` on the data and use the borders of the estimated `lambda` as upper and lower borders of the tuning space.

## Inspect fitted models during CV

Inspect Ridge regression on VI task in detail because the error is enourmus.

First extract the models.

```{r}
models <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-CV"]][["models"]]
```

Then look at the fold performances

```{r}
benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["measures.test"]][["rmse"]]
```

We see a high error on Fold 2 (= Laukiz 2).
The others are also quite high but not "out of bounds".

Because this models used the internal optimization of the lambda sequence (`cv.glmnet`), let's look at the value which was chosen for prediction (parameter `s` which defaults to `s="lambda.1se"`):

```{r}
purrr::map_dbl(models, ~ .x[["learner.model"]][["lambda.1se"]])
```

It seems that the `lambda.1se` value for Fold 4 is way higher than for the other 3 folds.
However, all values seem to be quite high.

Let's look at the full lambda sequence

```{r}
purrr::map_int(models, ~ length(.x[["learner.model"]][["lambda"]]))
```

Interestingly, the lambda length of fold 1 is not 100 (default) but only 5.

## Train/predict manually for Fold 1 via {glmnet}

To inspect further, let's refit a {glmnet} model directly on the training data of Fold 1 and inspect what `glmnet::cv.glmnet` estimates for the lambda sequence:

```{r}
train_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["train.inds"]][[4]]

obs_train_f4 <- as.matrix(task_new_buffer2[[2]]$env$data[train_inds_fold4, getTaskFeatureNames(task_new_buffer2[[2]])])
target_f4 <- getTaskTargets(task_new_buffer2[[2]])[train_inds_fold4]
```

Fit `cv.glmnet`

```{r}
set.seed(1)
modf4 <- glmnet::cv.glmnet(obs_train_f4, target_f4, alpha = 0)

modf4$lambda.1se
```

Ok, a value of `0.85` is **very** different to what happened during the CV (4.211054e+08).

Predict on Laukiz 2 now.

```{r}
pred_inds_fold4 <- benchmark_models_new_penalized_mbo_buffer2[[8]][["results"]][["vi_buffer2"]][["Ridge-MBO"]][["pred"]][["instance"]][["test.inds"]][[4]]

obs_pred_f4 <- as.matrix(task_new_buffer2[[2]]$env$data[pred_inds_fold4, getTaskFeatureNames(task_new_buffer2[[2]])])

pred <- predict(modf4, newx = obs_pred_f4, s = modf4$lambda.1se)
```

Calculate the error

```{r}
truth <- task_new_buffer2[[2]]$env$data[pred_inds_fold4, "defoliation"]

mlr:::measureRMSE(truth, pred)
```

Ok, RMSE of 97073324139.
This is most likely because of a few. observations which were predicted completely out of bounds.

```{r}
qplot(pred, geom = "histogram")
```

Ok, its one observation (row id  = 737).

Let's have a look at the predictor values for this observation.

```{r}
summary(obs_train_f4[737, ])
```

Ok, how does this compare to summaries of other observations?

```{r}
lapply(seq(500:510), function(x) summary(obs_train_f4[x, ]))
```

We have some higher values for obs 737 but nothing which stands out.

Let's look at the model coefficients and Partial Dependence Plots (PDP):

```{r}
coef(modf4)
```

Feature "bf2_PRI_norm" has a quite high value.

Let's figure out which are the ten most important features and create PDPs for these:

```{r}
top_ten_abs <- coef(modf4) %>%
  as.matrix() %>%
  as.data.frame() %>%
  dplyr::rename(coef = `1`) %>%
  dplyr::mutate(feature = rownames(coef(modf4))) %>%
  dplyr::slice(-1) %>%
  dplyr::mutate(coef_abs = abs(coef)) %>%
  dplyr::arrange(desc(coef_abs)) %>%
  dplyr::slice(1:10) %>%
  dplyr::pull(feature)
```

### Partial Dependence Plots

For PDP we use a model trained with {mlr} and check for equality first.

```{r}
lrn <- makeLearner("regr.cvglmnet", alpha = 0)
task_f4 <- subsetTask(task_new_buffer2[[2]], train_inds_fold4)

set.seed(1)
mod_mlr <- train(lrn, task_f4)
```

Check lambda sequence and `lambda.1se`:

```{r}
mod_mlr$learner.model$lambda
```

```{r}
mod_mlr$learner.model$lambda.1se
```

Check for equality between {mlr} and {glmnet} directly

```{r}
all.equal(modf4$lambda.1se, mod_mlr$learner.model$lambda.1se)
```

```{r}
pdp <- generatePartialDependenceData(mod_mlr, task_f4, features = top_ten_abs)
```

```{r}
plotPartialDependence(pdp)
```

Individual PDP

```{r}
pdp_ind <- generatePartialDependenceData(mod_mlr, task_f4,
  features = top_ten_abs,
  individual = TRUE
)
plotPartialDependence(pdp_ind)
```

## Compare arguments of the `do.call` call in the benchmark with the manual one

`args` were saved during a debug call of `mlr::benchmark()`

```{r}
args <- readRDS("args-bm-laukiz2.rda")
target_bm <- args$y
features_bm <- args$x

train_inds_bm <- as.numeric(rownames(features_bm))

all.equal(sort(target_bm), sort(target_f4))

all.equal(sort(train_inds_bm), sort(train_inds_fold4))
```