### 2020-04-18 docs/00-manuscripts/ieee/performance-svm-single-plot.tex

% latex table generated in R 3.6.1 by xtable 1.8-3 package
\begin{table}[ht!]
\centering
\caption{Test fold performances for learner SVM on the HR dataset without using a filter. For each row, the model was trained on observations from fold all other plots and tested on the observations of the given plot.} 
\label{tab:svm-single-fold-perf}
\begin{tabular}{rlr}
  \hline
 & Plot & RMSE \\ 
  \hline
1 & Luiando & 21.17 \\ 
  2 & Laukiz1 & 28.05 \\ 
  3 & Laukiz2 & 9.00 \\ 
  4 & Oiartzun & 54.26 \\ 
   \hline
\end{tabular}
\end{table}