### 2020-06-21 docs/00-manuscripts/ieee/performance-best-per-learner.tex
line-endings

% latex table generated in R 3.6.1 by xtable 1.8-3 package
\begin{table}[ht!]
\centering
\caption{Best performance of each learner across any task and filter method, sorted ascending by RMSE} 
\label{tab:best-learner-perf}
\scalebox{0.9}{
\begin{tabular}{rlllrr}
  \hline
 & Task & Model & Filter & RMSE & SE \\ 
  \hline
1 & NRI & SVM & Info Gain & 27.99 & 19.15 \\ 
  2 & NRI & RF & Car & 30.77 & 16.86 \\ 
  3 & VI & Lasso-MBO & No Filter & 31.01 & 14.71 \\ 
  4 & HR-NRI-VI & XGBoost & Borda & 31.05 & 17.01 \\ 
  5 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
   \hline
\end{tabular}
}
\end{table}