### 2020-12-25 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

---
title: "Exploratory data analysis"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

library("drake")
library("magrittr")
R.utils::sourceDirectory("R")
R.utils::sourceDirectory("code")

# load drake objects
loadd(
  vi_data_corrected_buffer2, nri_data_corrected_buffer2,
  bands_data_corrected_buffer2, coords_vi_nri_clean_corrected,

  filter_values,
  filter_info_gain_nbins,

  trees_with_bands_corrected_buffer2,
  trees_with_bands_corrected_no_buffer
)
library("DataExplorer")
library("dplyr")
library("ggsci")
library("ggpubr")
library("ggplot2")
library("knitr")
library("purrr")
library("sp")
library("raster")
library("fs")
```

VI = Vegetation Indices
NRI = Normalized Ratio Indices
HR = Hyperspectral reflectances (image bands)

# Datasets {.tabset .tabset-fade}

## VI

### Overview

```{r eda-1 }
intro <- introduce(vi_data_corrected_buffer2)
intro_df <- data.frame(
  "Name" = c(
    "Rows", "Columns",
    "Discrete columns", "Continuous columns", "All missing columns",
    "Missing observations", "Complete Rows",
    "Total observations"
  ),
  "Value" = c(
    format(intro[["rows"]], big.mark = ","),
    format(intro[["columns"]], big.mark = ","),
    format(intro[["discrete_columns"]], big.mark = ","),
    format(intro[["continuous_columns"]], big.mark = ","),
    format(intro[["all_missing_columns"]], big.mark = ","),
    format(intro[["total_missing_values"]], big.mark = ","),
    format(intro[["complete_rows"]], big.mark = ","),
    format(intro[["total_observations"]], big.mark = ",")
  )
)
knitr::kable(intro_df)
```

### Histograms

```{r eda-2 }
plot_histogram(vi_data_corrected_buffer2)
```

### PCA

```{r eda-3 }
vi_data_corrected_buffer2_no_defol = vi_data_corrected_buffer2
vi_data_corrected_buffer2_no_defol$defoliation <- NULL
pca <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol,
  variance_cap = 0.98,
  ggtheme = ggpubr::theme_pubclean()
)
pca[[1]]
```

### Corr

```{r eda-4 }
plot_correlation(vi_data_corrected_buffer2)
```

## NRI

### Overview

```{r eda-5 }
# remove response
nri_data_corrected_buffer2$defoliation <- NULL

intro <- introduce(nri_data_corrected_buffer2)
intro_df <- data.frame(
  "Name" = c(
    "Rows", "Columns",
    "Discrete columns", "Continuous columns", "All missing columns",
    "Missing observations", "Complete Rows",
    "Total observations"
  ),
  "Value" = c(
    format(intro[["rows"]], big.mark = ","),
    format(intro[["columns"]], big.mark = ","),
    format(intro[["discrete_columns"]], big.mark = ","),
    format(intro[["continuous_columns"]], big.mark = ","),
    format(intro[["all_missing_columns"]], big.mark = ","),
    format(intro[["total_missing_values"]], big.mark = ","),
    format(intro[["complete_rows"]], big.mark = ","),
    format(intro[["total_observations"]], big.mark = ",")
  )
)
kable(intro_df)
```

### Histograms

No histograms for NRI -> too many features.

### PCA

```{r eda-6 }
pca <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2,
  variance_cap = 0.98,
  ggtheme = ggpubr::theme_pubclean()
)
pca[[1]]
```

### Corr

```{r eda-7 }
plot_correlation(nri_data_corrected_buffer2)
```

## HR

### Overview

```{r eda-8 }
# remove response
bands_data_corrected_buffer2$defoliation <- NULL

intro <- introduce(bands_data_corrected_buffer2)
intro_df <- data.frame(
  "Name" = c(
    "Rows", "Columns",
    "Discrete columns", "Continuous columns", "All missing columns",
    "Missing observations", "Complete Rows",
    "Total observations"
  ),
  "Value" = c(
    format(intro[["rows"]], big.mark = ","),
    format(intro[["columns"]], big.mark = ","),
    format(intro[["discrete_columns"]], big.mark = ","),
    format(intro[["continuous_columns"]], big.mark = ","),
    format(intro[["all_missing_columns"]], big.mark = ","),
    format(intro[["total_missing_values"]], big.mark = ","),
    format(intro[["complete_rows"]], big.mark = ","),
    format(intro[["total_observations"]], big.mark = ",")
  )
)
kable(intro_df)
```

### Histograms

```{r eda-9 }
plot_histogram(bands_data_corrected_buffer2)
```

### PCA

```{r eda-10 pca-bands, dev = c("png", "pdf")}
pca <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2,
  variance_cap = 0.98,
  ggtheme = ggpubr::theme_pubclean()
)
pca[[1]]
```

### Corr

```{r eda-11 }
plot_correlation(bands_data_corrected_buffer2)
```

# Custom plots

```{r eda-12 }
vi_data_plot <- vi_data_corrected_buffer2 %>%
  mutate(plot = factor(rep(
    c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
    c(559, 451, 301, 497)
  )))
```

# Mean defoliation per plot

```{r eda-13 }
mean_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(mean(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(mean(defoliation))
```

# Coefficient of variation

```{r eda-14 }
cov_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise((sd(defoliation) / mean(defoliation)) * 100) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise((sd(defoliation) / mean(defoliation)) * 100)
```

# sd / skewness

```{r eda-15 }
sd_skewness_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation))
```

```{r eda-16 }
sd_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(sd(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(sd(defoliation))
```

```{r defoliation-distribution-plot, dev = c("png", "pdf")}
boxplot_defol <- vi_data_plot %>%
  group_by(plot) %>%
  ggboxplot(
    x = "plot", y = "defoliation", color = "plot",
    add = "jitter", add.params = list(size = "defoliation")
  ) +
  scale_size(range = c(0.5, 0.5)) +
  annotate("text",
    label = expression(bold(atop("n = 559", bar(x) ~ "= 57.23"))), x = 1,
    y = 112, size = 4, colour = "#BC3C29", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 451", bar(x) ~ "= 13.54"))), x = 2,
    y = 112, size = 4, colour = "#0072B5", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 301", bar(x) ~ "= 68.36"))), x = 3,
    y = 112, size = 4, colour = "#E18727", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 497", bar(x) ~ "= 69.07"))), x = 4,
    y = 112, size = 4, colour = "#20854E", fontface = 2
  ) +
  scale_color_nejm() +
  ggpubr::theme_pubr(base_size = 14) +
  theme(legend.position = "none") +
  labs(y = "Total defoliation per tree (%)", x = "Plot")
boxplot_defol
```

# Point density

In Meters.

```{r eda-18 }
plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
dist <- map(plots, ~ {
  coords <- coords_vi_nri_clean_corrected %>%
    mutate(plot = factor(rep(
      c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
      c(559, 451, 301, 497)
    ))) %>%
    filter(plot == .x) %>%
    dplyr::select(-plot)
  points <- SpatialPoints(
    coords = coords,
    proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
  )
  distance <- mean(as.dist(pointDistance(points,
    allpairs = TRUE,
    lonlat = FALSE
  )))
})
set_names(dist, plots)
dist_plots <- unlist(dist)
```

```{r eda-19, echo = FALSE}
# file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
#           "code/98-paper/journal/")
```

# Effects of different buffer sizes when extracting values to trees

The following buffer sizes were used to extract the remote sensing information (1 m spatial resolution) to the trees (points):

- 0 m
- 0.5 m
- 1 m
- 1.5 m
- 2 m

For this EDA only the direct reflectance values are used.
It is assumed that the VI and NRI datasets will be affected in the same way by the different buffer sizes.

```{r buffer-size-comp}
compare_buffer <- function(trees_with_buffer, trees_without_buffer, plot_name) {
  if (plot_name == "laukiz1") {
    targets_sub_ids <- 1
  } else if (plot_name == "laukiz2") {
    targets_sub_ids <- 2
  } else if (plot_name == "luiando") {
    targets_sub_ids <- 3
  } else if (plot_name == "oiartzun") {
    targets_sub_ids <- 4
  }

  foo_buffer <- trees_with_buffer[seq(targets_sub_ids, 12, 4)]

  foo_buffer <- foo_buffer %>%
    map(~ sf::st_set_geometry(., NULL)) %>%
    map(~ dplyr::summarise_at(., vars(starts_with("B")), mean, na.rm = TRUE)) %>%
    map(~ tidyr::pivot_longer(., everything(),
      names_to = "band", values_to = "reflectance"
    )) %>%
    dplyr::bind_rows(.id = "Buffer size")

  foo_no_buffer <- trees_without_buffer[[targets_sub_ids]]

  foo_no_buffer <- foo_no_buffer %>%
    sf::st_set_geometry(NULL) %>%
    dplyr::summarise_at(vars(starts_with("B")), mean) %>%
    tidyr::pivot_longer(everything(),
      names_to = "band", values_to = "reflectance"
    ) %>%
    dplyr::bind_rows(.id = "Buffer size") %>%
    dplyr::mutate(`Buffer size` = "No Buffer")

  all <- bind_rows(foo_buffer, foo_no_buffer) %>%
    dplyr::mutate(band = as.numeric(stringr::str_replace(band, "B", "")))

  plot <- ggplot(all, aes(x = band, y = reflectance)) +
    geom_line(aes(color = `Buffer size`), size = 0.6) +
    labs(title = tools::toTitleCase(plot_name)) +
    ggsci::scale_color_nejm() +
    ggpubr::theme_pubr(legend = "right")

  return(plot)
}
```

```{r}
map(
  c("laukiz1", "laukiz2", "luiando", "oiartzun"),
  ~ compare_buffer(
    trees_with_bands_corrected_buffer2,
    trees_with_bands_corrected_no_buffer,
    .x
  )
) %>%
  patchwork::wrap_plots() +
  patchwork::guide_area() +
  patchwork::plot_layout(guides = "collect", ncol = 3, nrow = 2) +
  patchwork::plot_annotation(
    title = "Comparing the effect of different buffer sizes when extracting reflectance values to points obsverations (trees)",
    subtitle = "Mean values of each hyperspectral band for each plot "
  )
```