### 2020-12-26 analysis/eda.Rmd
wflow_publish(knitr_in("analysis/eda.Rmd"), view = FALSE, verbose = TRUE)

---
title: "Exploratory data analysis"
output: 
  workflowr::wflow_html:
    includes:
      in_header: header.html
editor_options:
  chunk_output_type: console
author: "Patrick Schratz"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 3,
  fig.align = "center",
  fig.width = 6.93,
  fig.height = 6.13,
  out.width = "100%",
  echo = FALSE
)

library("drake")
library("magrittr")
R.utils::sourceDirectory("R")
R.utils::sourceDirectory("code")

# load drake objects
loadd(
  vi_data_corrected_buffer2, nri_data_corrected_buffer2,
  bands_data_corrected_buffer2, coords_vi_nri_clean_corrected,

  filter_values,
  filter_info_gain_nbins,

  trees_with_bands_corrected_buffer2,
  trees_with_bands_corrected_no_buffer
)
library("DataExplorer")
library("dplyr")
library("ggsci")
library("ggpubr")
library("ggplot2")
library("knitr")
library("purrr")
library("sp")
library("raster")
library("fs")
```

- VI = Vegetation Indices
- NRI = Normalized Ratio Indices
- HR = Hyperspectral reflectances (image bands)

# Datasets {.tabset .tabset-fade}

## VI

### Overview

```{r eda-1 }
intro <- introduce(vi_data_corrected_buffer2)
intro_df <- data.frame(
  "Name" = c(
    "Rows", "Columns",
    "Discrete columns", "Continuous columns", "All missing columns",
    "Missing observations", "Complete Rows",
    "Total observations"
  ),
  "Value" = c(
    format(intro[["rows"]], big.mark = ","),
    format(intro[["columns"]], big.mark = ","),
    format(intro[["discrete_columns"]], big.mark = ","),
    format(intro[["continuous_columns"]], big.mark = ","),
    format(intro[["all_missing_columns"]], big.mark = ","),
    format(intro[["total_missing_values"]], big.mark = ","),
    format(intro[["complete_rows"]], big.mark = ","),
    format(intro[["total_observations"]], big.mark = ",")
  )
)
knitr::kable(intro_df)
```

### Histograms

```{r eda-2 }
plot_histogram(vi_data_corrected_buffer2)
```

### PCA

```{r eda-3 }
# remove response
vi_data_corrected_buffer2_no_defol_no_response <- vi_data_corrected_buffer2
vi_data_corrected_buffer2_no_defol_no_response$defoliation <- NULL

pca_vi <- DataExplorer::plot_prcomp(vi_data_corrected_buffer2_no_defol_no_response,
  variance_cap = 0.90,
  ggtheme = ggpubr::theme_pubclean(),
  geom_label_args = list(size = 3),
  parallel = TRUE
)
pca_vi[[1]]
```

### Corr

```{r eda-4 }
plot_correlation(vi_data_corrected_buffer2)
```

## NRI

### Overview

```{r eda-5 }
intro <- introduce(nri_data_corrected_buffer2)
intro_df <- data.frame(
  "Name" = c(
    "Rows", "Columns",
    "Discrete columns", "Continuous columns", "All missing columns",
    "Missing observations", "Complete Rows",
    "Total observations"
  ),
  "Value" = c(
    format(intro[["rows"]], big.mark = ","),
    format(intro[["columns"]], big.mark = ","),
    format(intro[["discrete_columns"]], big.mark = ","),
    format(intro[["continuous_columns"]], big.mark = ","),
    format(intro[["all_missing_columns"]], big.mark = ","),
    format(intro[["total_missing_values"]], big.mark = ","),
    format(intro[["complete_rows"]], big.mark = ","),
    format(intro[["total_observations"]], big.mark = ",")
  )
)
kable(intro_df)
```

### Histograms

No histograms for NRI -> too many features.

### PCA

```{r eda-6 }
# remove response
nri_data_corrected_buffer2_no_response <- nri_data_corrected_buffer2
nri_data_corrected_buffer2_no_response$defoliation <- NULL

pca_nri <- DataExplorer::plot_prcomp(nri_data_corrected_buffer2_no_response,
  variance_cap = 0.9,
  ggtheme = ggpubr::theme_pubclean(),
  geom_label_args = list(size = 3),
  parallel = TRUE
)
pca_nri[[1]]
```

### Corr

No correlation plot for NRI -> too many features.

## HR

### Overview

```{r eda-8 }
intro <- introduce(bands_data_corrected_buffer2)
intro_df <- data.frame(
  "Name" = c(
    "Rows", "Columns",
    "Discrete columns", "Continuous columns", "All missing columns",
    "Missing observations", "Complete Rows",
    "Total observations"
  ),
  "Value" = c(
    format(intro[["rows"]], big.mark = ","),
    format(intro[["columns"]], big.mark = ","),
    format(intro[["discrete_columns"]], big.mark = ","),
    format(intro[["continuous_columns"]], big.mark = ","),
    format(intro[["all_missing_columns"]], big.mark = ","),
    format(intro[["total_missing_values"]], big.mark = ","),
    format(intro[["complete_rows"]], big.mark = ","),
    format(intro[["total_observations"]], big.mark = ",")
  )
)
kable(intro_df)
```

### Histograms

```{r eda-9 }
plot_histogram(bands_data_corrected_buffer2)
```

### PCA

```{r eda-10 }
# remove response
bands_data_corrected_buffer2_no_response <- bands_data_corrected_buffer2
bands_data_corrected_buffer2_no_response$defoliation <- NULL

pca_hr <- DataExplorer::plot_prcomp(bands_data_corrected_buffer2_no_response,
  variance_cap = 0.90,
  geom_label_args = list(size = 3),
  ggtheme = ggpubr::theme_pubclean()
)
pca_hr[[1]]
```

### Corr

```{r eda-11 }
plot_correlation(bands_data_corrected_buffer2)
```

# Custom plots

```{r eda-12 }
vi_data_plot <- vi_data_corrected_buffer2 %>%
  mutate(plot = factor(rep(
    c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
    c(559, 451, 301, 497)
  )))
```

# Mean defoliation per plot

```{r eda-13 }
mean_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(mean(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(mean(defoliation))
```

# Coefficient of variation

```{r eda-14 }
cov_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise((sd(defoliation) / mean(defoliation)) * 100) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise((sd(defoliation) / mean(defoliation)) * 100)
```

# sd / skewness

```{r eda-15 }
sd_skewness_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(((sd(defoliation) / mean(defoliation)) * 100) / e1071::skewness(defoliation))
```

```{r eda-16 }
sd_defol <- vi_data_plot %>%
  group_by(plot) %>%
  summarise(sd(defoliation)) %>%
  pull(.)
vi_data_plot %>%
  group_by(plot) %>%
  summarise(sd(defoliation))
```

```{r defoliation-distribution-plot, dev = c("png", "pdf")}
boxplot_defol <- vi_data_plot %>%
  group_by(plot) %>%
  ggboxplot(
    x = "plot", y = "defoliation", color = "plot",
    add = "jitter", add.params = list(size = "defoliation")
  ) +
  scale_size(range = c(0.5, 0.5)) +
  annotate("text",
    label = expression(bold(atop("n = 559", bar(x) ~ "= 57.23"))), x = 1,
    y = 112, size = 4, colour = "#BC3C29", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 451", bar(x) ~ "= 13.54"))), x = 2,
    y = 112, size = 4, colour = "#0072B5", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 301", bar(x) ~ "= 68.36"))), x = 3,
    y = 112, size = 4, colour = "#E18727", fontface = 2
  ) +
  annotate("text",
    label = expression(bold(atop("n = 497", bar(x) ~ "= 69.07"))), x = 4,
    y = 112, size = 4, colour = "#20854E", fontface = 2
  ) +
  scale_color_nejm() +
  ggpubr::theme_pubr(base_size = 14) +
  theme(legend.position = "none") +
  labs(y = "Total defoliation per tree (%)", x = "Plot")
boxplot_defol
```

# Point density

In Meters.

```{r eda-18 }
plots <- list("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun")
dist <- map(plots, ~ {
  coords <- coords_vi_nri_clean_corrected %>%
    mutate(plot = factor(rep(
      c("Laukiz 1", "Laukiz 2", "Luiando", "Oiartzun"),
      c(559, 451, 301, 497)
    ))) %>%
    filter(plot == .x) %>%
    dplyr::select(-plot)
  points <- SpatialPoints(
    coords = coords,
    proj4string = CRS("+proj=utm +zone=30 +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
  )
  distance <- mean(as.dist(pointDistance(points,
    allpairs = TRUE,
    lonlat = FALSE
  )))
})
set_names(dist, plots)
dist_plots <- unlist(dist)
```

```{r eda-19, echo = FALSE}
# file_move("docs/figure/eda.Rmd/defoliation-distribution-plot-1.pdf",
#           "code/98-paper/journal/")
# file_move("docs/figure/eda.Rmd/pca-bands-1.pdf",
# "code/98-paper/journal/")
```

# Effects of different buffer sizes when extracting values to trees

The following buffer sizes were used to extract the remote sensing information (1 m spatial resolution) to the trees (points):

- 0 m
- 0.5 m
- 1 m
- 1.5 m
- 2 m

For this EDA only the direct reflectance values are used.
It is assumed that the VI and NRI datasets will be affected in the same way by the different buffer sizes.

```{r buffer-size-comp}
compare_buffer <- function(trees_with_buffer, trees_without_buffer, plot_name) {
  if (plot_name == "laukiz1") {
    targets_sub_ids <- 1
  } else if (plot_name == "laukiz2") {
    targets_sub_ids <- 2
  } else if (plot_name == "luiando") {
    targets_sub_ids <- 3
  } else if (plot_name == "oiartzun") {
    targets_sub_ids <- 4
  }

  foo_buffer <- trees_with_buffer[seq(targets_sub_ids, 12, 4)]

  foo_buffer <- foo_buffer %>%
    map(~ sf::st_set_geometry(., NULL)) %>%
    map(~ dplyr::summarise_at(., vars(starts_with("B")), mean, na.rm = TRUE)) %>%
    map(~ tidyr::pivot_longer(., everything(),
      names_to = "band", values_to = "reflectance"
    )) %>%
    dplyr::bind_rows(.id = "Buffer size")

  foo_no_buffer <- trees_without_buffer[[targets_sub_ids]]

  foo_no_buffer <- foo_no_buffer %>%
    sf::st_set_geometry(NULL) %>%
    dplyr::summarise_at(vars(starts_with("B")), mean) %>%
    tidyr::pivot_longer(everything(),
      names_to = "band", values_to = "reflectance"
    ) %>%
    dplyr::bind_rows(.id = "Buffer size") %>%
    dplyr::mutate(`Buffer size` = "No Buffer")

  all <- bind_rows(foo_buffer, foo_no_buffer) %>%
    dplyr::mutate(band = as.numeric(stringr::str_replace(band, "B", "")))

  plot <- ggplot(all, aes(x = band, y = reflectance)) +
    geom_line(aes(color = `Buffer size`), size = 0.6) +
    labs(title = tools::toTitleCase(plot_name)) +
    ggsci::scale_color_nejm() +
    ggpubr::theme_pubr(legend = "right")

  return(plot)
}
```

```{r buffer-size-comp-plot}
map(
  c("laukiz1", "laukiz2", "luiando", "oiartzun"),
  ~ compare_buffer(
    trees_with_bands_corrected_buffer2,
    trees_with_bands_corrected_no_buffer,
    .x
  )
) %>%
  patchwork::wrap_plots() +
  patchwork::guide_area() +
  patchwork::plot_layout(guides = "collect", ncol = 3, nrow = 2) +
  patchwork::plot_annotation(
    title = "Comparing the effect of different buffer sizes when extracting reflectance values to points obsverations (trees)",
    subtitle = "Mean values of each hyperspectral band for each plot "
  )
```

# PCA feature set comparison

Comparing the "proportion of variance explained" (POV) for all feature sets.

```{r pca-pov-comparison, dev = c("png", "pdf")}
pca_vi[[1]] / pca_nri[[1]] / pca_hr[[1]] *
  labs(title = NULL) *
  theme(
    axis.title.y = element_text(size = 8),
    axis.title.x = element_text(size = 8)
  ) +
  patchwork::plot_annotation(
    tag_levels = list(c("VI", "NRI", "HR")),
    title = "Proportin of variance explained by principal components",
    subtitle = "Labels indicate cumulative % of explained variance"
  )
```