### 2021-02-19 analysis/about.Rmd
workflowr::wflow_publish("analysis/about.Rmd")

---
title: "About"
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

**Modeling defoliation as a proxy for tree health: Comparison of feature-selection methods across multiple feature sets derived from hyperspectral data**

# Authors

**Patrick Schratz** (patrick.schratz@gmail.com) [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0003-0748-6624)

Jannes Muenchow [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-7834-4717)

Eugenia Iturritxa [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](https://orcid.org/0000-0002-0577-3315)

José Cortés [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0003-2567-8689)

Bernd Bischl [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6002-6980)

Alexander Brenning [![](https://orcid.org/sites/default/files/images/orcid_16x16.png)](http://orcid.org/0000-0001-6640-679X)

# Contents

## Paper

This repository contains the research compendium of our work on comparing algorithms across multiple feature sets and filtering methods (including ensemble filter methods).

- keywords

  - hyperspectral imagery
  - forest health monitoring
  - machine learning
  - feature selection
  - feature effects
  - model comparison
  - filter
  - imaging spectroscopy

- Using machine-learning algorithms to model defoliation of _Pinus Radiata_ trees.

- Compare filtering methods (ensemble filter methods) across various algorithms and datasets

- Predict defoliation to all available plots (24) and the whole Basque Country (at 200 m resolution)

The following directories belong to this project

- `code/01-download.R`
- `code/02-hyperspectral-processing.R`
- `code/04-data-processing.R`
- `code/05-modeling/`
- `code/06-benchmark-matrix/`
- `code/07-reports/`

## Other Content

In addition, this repo contains the workflow for an analysis related to the [LIFE 14 ENV/ES/000179 LIFE HEALTHY FOREST](http://www.lifehealthyforest.com/) project: Predicting defoliation at trees for the Basque Country (for the years 2017 and 2018) using Sentinel-2 data.

Target `defoliation_maps_wfr` builds subsequent argets which are necessary for the final results [report](https://pat-s.github.io/2019-feature-selection/report-defoliation.html).

# How to use

## Reading the code, accessing the data

See the [`code`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis) directory on GitHub for the source code that generated the figures and statistical results contained in the manuscript.
See the [`data`](https://github.com/pat-s/paper_hyperspectral/tree/master/analysis/data) directory for instructions how to access the raw data used in the manuscript.

## Installing the R package

This repository is organized as an R package, providing functions and raw data to reproduce and extend the analysis reported in the publication.
Note that this package has been written explicitly for this project and is not suited a for more general use.

This project is setup with a [drake workflow](https://github.com/ropensci/drake), ensuring reproducibility.
Intermediate targets/objects will be stored in a hidden `.drake` directory.

The R library of this project is managed by [renv](https://rstudio.github.io/renv/index.html).
This makes sure that the exact same package versions are used when recreating the project.
When calling `renv::restore()`, all required packages will be installed with their specific version.

Please note that this project was built with R version 4.0.3 on a CentOS 7 operating system.
Some packages from this project **are not compatible with R versions prior version 3.6.0.**

To clone the project, a working installation of `git` is required.
Open a terminal in the directory of your choice and execute:

```sh
git clone https://github.com/pat-s/2019-feature-selection.git
```

Then start R in this directory and run

```{r README-1, eval = FALSE}
renv::restore()
r_make()
```

## Creating targets with {drake}

Calling `r_make()` will create targets specified in `drake_config(targets = <target>)` in `_drake.R` with the additional drake settings specified.

Out of the 400+ targets in this project, the following targets are important:

-   `bm_aggregated_new_buffer2`: Aggregated benchmark results of all models using a 2 meter buffer for hyperspectral data extraction.
-   `eda_wfr`: Creates the [report which shows Exploratory Data Analysis (EDA) plots and tables](https://pat-s.github.io/2019-feature-selection/eda.html).
-   `eval_performance_wfr`: Creates the [report which evaluates the model performances](https://pat-s.github.io/2019-feature-selection/eval-performance.html).
-   `spectral_signatures_wfr`: Creates the [report which inspects the spectral signatures of the hyperspectral data](https://pat-s.github.io/2019-feature-selection/spectral-signatures.html).
-   `feature_importance_wfr`: Creates the [report which inspects the feature importance of variables](https://pat-s.github.io/2019-feature-selection/feature-importance.html).
-   `filter_correlations_wfr`: Creates the [report which inspects correlations among filter methods](https://pat-s.github.io/2019-feature-selection/feature-importance.html).

Note that most reports require some/all fitted models.
Creating these (e.g. target `benchmark_no_models_new_buffer2`) is a costly process and takes several days on a HPC and way longer on a single machine.

# Notes and resources

* The organisation of this compendium was inspired by the works of [Carl Boettiger](http://www.carlboettiger.info/) and [Ben Marwick](https://github.com/benmarwick).