### 2021-02-19 analysis/index.Rmd
workflowr::wflow_publish("analysis/index.Rmd")

---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

[![minimal R version](https://img.shields.io/badge/R%3E%3D-4.0.3-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2582969.svg)](https://doi.org/10.5281/zenodo.2582969)

This [workflowr]() site contains all information and results about this research project.

Project details can be found in the [About](about.html) page.

**Result shortcuts**

1. [Prediction maps of defoliation at trees](report-defoliation.html)

Powered by [drake](https://github.com/ropensci/drake) + [workflowr](https://github.com/jdblischak/workflowr) + [mlr](https://github.com/mlr-org/mlr).

# Acknowledgements

This work has been supported by the the European Union "Healthy Forest" Project LIFE14 ENV/ES/000179.