### 2021-02-24 docs/00-manuscripts/ieee/performance-best-per-learner.tex
update latex performance tables

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Best performance of each learner across any task and filter method, sorted ascending by RMSE} 
\label{tab:best-learner-perf}
\scalebox{0.9}{
\begin{tabular}{rlllrrr}
  \hline
 & Task & Model & Filter & rsq & RMSE & SE \\ 
  \hline
1 & NRI & SVM & Info Gain & -0.07 & 28.03 & 19.13 \\ 
  2 & HR-NRI & XGBoost & Car & -0.30 & 31.02 & 17.63 \\ 
  3 & VI & Lasso-MBO & No Filter & -5.97 & 31.11 & 14.91 \\ 
  4 & NRI & Ridge-MBO & No Filter & -0.18 & 31.16 & 15.03 \\ 
  5 & NRI & RF & Car & -5.97 & 32.14 & 15.97 \\ 
   \hline
\end{tabular}
}
\end{table}