### 2021-04-09 docs/00-manuscripts/ieee/performance-svm-single-plot.tex

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Test fold performances for learner SVM on the HR dataset without using a filter. For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot.} 
\label{tab:svm-single-fold-perf}
\begin{tabular}{rrrl}
  \hline
 & Plot & RMSE & Test Plot \\ 
  \hline
1 &   1 & 28.12 & Laukiz1 \\ 
  2 &   2 & 54.26 & Laukiz2 \\ 
  3 &   3 & 9.00 & Luiando \\ 
  4 &   4 & 21.17 & Oiartzun \\ 
   \hline
\end{tabular}
\end{table}