### 2021-04-09 docs/00-manuscripts/ieee/tune-perc-sel-features.tex

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Selected feature portions during tuning for selected learner-filter settings across folds for task HR-NRI-VI, sorted ascending by RMSE} 
\label{tab:tune-perc-sel-features}
\begin{tabular}{llrrr}
  \toprule
Learner & Test Plot & Features (\%) & \# & RMSE \\ 
  \midrule
\multirow{4}{*}{\specialcell{RF \\ Car}} & Laukiz1 & 0.14212 & 69 & 41.48 \\ 
   & Laukiz2 & 0.21934 & 99 & 16.47 \\ 
   & Luiando & 0.02878 & 9 & 46.33 \\ 
   & Oiartzun & 0.00279 & 2 & 31.82 \\ 
  \midrule\multirow{4}{*}{\specialcell{XGB \\ Borda}} & Laukiz1 & 0.01502 & 8 & 15.48 \\ 
   & Laukiz2 & 0.66619 & 301 & 44.39 \\ 
   & Luiando & 0.10428 & 32 & 40.86 \\ 
   & Oiartzun & 0.66651 & 353 & 34.15 \\ 
  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
   & Laukiz2 & 0.04106 & 19 & 31.03 \\ 
   & Luiando & 0.20735 & 63 & 37.19 \\ 
   & Oiartzun & 0.23928 & 127 & 14.89 \\ 
   \bottomrule
\end{tabular}
\end{table}