### 2021-04-12 docs/00-manuscripts/ieee/tune-perc-sel-features.tex

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Selected feature portions during tuning for the best performing learner-filter settings (SVM Relief, RF Relief, XGBoost CMIM) across folds for task HR-NRI-VI, sorted ascending by RMSE.} 
\label{tab:tune-perc-sel-features}
\begin{tabular}{llrrr}
  \toprule
Learner & Test Plot & Features (\%) & \# & RMSE \\ 
  \midrule
\multirow{4}{*}{\specialcell{RF \\ Relief}} & Laukiz1 & 0.99973 & 559 & 41.73 \\ 
   & Laukiz2 & 0.93371 & 422 & 17.12 \\ 
   & Luiando & 0.00216 & 1 & 33.21 \\ 
   & Oiartzun & 0.00004 & 1 & 37.20 \\ 
  \midrule\multirow{4}{*}{\specialcell{XGB \\ CMIM}} & Laukiz1 & 0.30849 & 173 & 33.36 \\ 
   & Laukiz2 & 0.98122 & 443 & 45.59 \\ 
   & Luiando & 0.83613 & 252 & 41.01 \\ 
   & Oiartzun & 0.00515 & 3 & 15.24 \\ 
  \midrule\multirow{4}{*}{\specialcell{SVM \\ Relief}} & Laukiz1 & 0.00062 & 1 & 35.49 \\ 
   & Laukiz2 & 0.04106 & 19 & 31.03 \\ 
   & Luiando & 0.20735 & 63 & 37.19 \\ 
   & Oiartzun & 0.23928 & 119 & 14.89 \\ 
   \bottomrule
\end{tabular}
\end{table}