### 2021-04-29 docs/00-manuscripts/ieee/performance-best-per-learner.tex

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Best performance of each learner across any task and filter method, sorted ascending by RMSE} 
\label{tab:best-learner-perf}
\scalebox{0.9}{
\begin{tabular}{rlllrr}
  \hline
 & Task & Model & Filter & RMSE & SE \\ 
  \hline
1 & HR-NRI-VI & SVM & Relief & 28.09 & 18.95 \\ 
  2 & HR & XGBoost & CMIM & 30.88 & 14.47 \\ 
  3 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
  4 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
  5 & HR & RF & Relief & 32.93 & 13.17 \\ 
   \hline
\end{tabular}
}
\end{table}