### 2021-09-02 docs/00-manuscripts/mdpi/performance-best-per-learner.tex
update table caption

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run. For 	exttt{regr.featureless} the Task is no applicable and was therefore removed.} 
\label{tab:best-learner-perf}
\scalebox{0.9}{
\begin{tabular}{rlllrr}
  \hline
 & Task & Model & Filter & RMSE & SE \\ 
  \hline
1 & HR & SVM & Relief & 28.12 & 19.12 \\ 
  2 & NRI & Lasso-MBO & No Filter & 31.16 & 15.03 \\ 
  3 & NRI & Ridge-MBO & No Filter & 31.16 & 15.03 \\ 
  4 & - & regr.featureless & No Filter & 31.16 & 15.03 \\ 
  5 & HR & RF & Relief & 31.25 & 14.96 \\ 
  6 & HR & XGBoost & Car & 34.07 & 13.57 \\ 
   \hline
\end{tabular}
}
\end{table}