### 2021-11-06 docs/00-manuscripts/mdpi/performance-svm-single-plot.tex
chmod 644 -> 755

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{Test fold performances in RMSE (p.p.) for learner SVM on the HR dataset without using a filter, showcasing performance variance on the fold level. For each row, the model was trained on observations from all others plots but the given one and tested on the observations of the given plot.} 
\label{tab:svm-single-fold-perf}
\begin{tabular}{rrl}
  \hline
 & RMSE & Test Plot \\ 
  \hline
1 & 54.26 & Laukiz1 \\ 
  2 & 28.12 & Laukiz2 \\ 
  3 & 9.00 & Luiando \\ 
  4 & 21.17 & Oiartzun \\ 
   \hline
\end{tabular}
\end{table}