### 2021-11-12 docs/00-manuscripts/mdpi/performance-best-per-learner.tex
ensure that only 122 instead uf 126 bands are used for all feature sets

% latex table generated in R 4.0.4 by xtable 1.8-4 package
\begin{table}[ht!]
\centering
\caption{The overall best individual learner performance across any task and filter method for RF, SVM, XGBoost, Lasso and Ridge, sorted ascending by RMSE (p.p.) including the respective standard error (SE) of the cross-validation run. For 	exttt{regr.featureless} the Task is no applicable and was therefore removed.} 
\label{tab:best-learner-perf}
\scalebox{0.9}{
\begin{tabular}{rlllrr}
  \hline
 & Task & Model & Filter & RMSE & SE \\ 
  \hline
1 & NRI-VI & SVM & Info Gain & 27.915 & 18.970 \\ 
  2 & NRI & RF & Relief & 30.842 & 12.028 \\ 
  3 & HR & XGBoost & Info Gain & 31.165 & 15.025 \\ 
  4 & NRI & Lasso-MBO & No Filter & 31.165 & 15.025 \\ 
  5 & NRI & Ridge-MBO & No Filter & 31.165 & 15.025 \\ 
  6 & - & regr.featureless & No Filter & 31.165 & 15.025 \\ 
   \hline
\end{tabular}
}
\end{table}